% =========================================================
% Bounding: Bounds, Suprema, Completeness — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ \textbf{Bounding (Bounds \& Completeness)}
$\;\to\;$ Convergence (Sequences)
$\;\to\;$ Distance (Metric Spaces)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Volume~II established $\mathbb{R}$ as a complete ordered field and
introduced the field and order axioms as the static control layer
of analysis.
Those axioms allow us to compare and combine quantities, but say
nothing about infinite processes.
Bounding is the first chapter that confronts the infinite directly:
it asks what it means for a set to be restrained from above or below,
and what guarantee $\mathbb{R}$ provides when such restraint exists.

\medskip
\noindent\textbf{What this chapter builds.}
Upper and lower bounds, the supremum and infimum as least and greatest
bounds, the completeness axiom (every nonempty set bounded above has a
supremum), the $\varepsilon$-characterization of suprema, uniqueness,
the Nested Interval Property, the Archimedean Property, the density of
$\mathbb{Q}$ in $\mathbb{R}$, and the existence of square roots.

\medskip
\noindent\textbf{Where this leads.}
Convergence requires bounding: a sequence cannot converge without
first being controlled.
Every $\varepsilon$-$N$ argument ultimately appeals to completeness
to guarantee that the limit exists.
The supremum property is also the engine behind the Monotone
Convergence Theorem, Bolzano--Weierstrass, and Heine--Borel.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Role
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Bounding — The Vocabulary of Restraint}},
  fonttitle=\small\bfseries
]
The ordered field axioms allow comparison and combination,
but they do not yet control infinite processes.

Bounding introduces the vocabulary of restraint.
Upper bounds, lower bounds, suprema, and infima are the first tools
that let us reason about sets of numbers without enumerating every
element.

The completeness axiom then asserts something deeper:
whenever values approach a boundary from below,
that boundary already exists inside $\mathbb{R}$.

This is not merely a technical condition.
It is a statement about infinite approximation:
$\mathbb{R}$ is closed under its own limiting behavior.

The supremum property, the nested interval principle, and the
Archimedean property are different expressions of the same structural
phenomenon.
They translate the static vocabulary of bounds into the first
guarantee that infinite approximation cannot escape.

All later limit and convergence arguments rest on this chapter.
Bounding is the language; completeness is the law.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Logical Structure}
\end{center}

The logical progression is ordered by increasing complexity:
bounds (universal quantifier only) precede suprema
(universal + approximation), which precede completeness
(existence of limits).

\begin{enumerate}
  \item Upper and lower bounds: predicate and existence forms
  \item Maximum and minimum: bounds that belong to the set
  \item Supremum (least upper bound) and infimum (greatest lower bound)
  \item Uniqueness of supremum and infimum
  \item $\varepsilon$-characterization of supremum and infimum
  \item Completeness Axiom: every nonempty bounded-above set has a supremum
  \item Nested Interval Property
  \item Archimedean Property and the integer part function
  \item Density of $\mathbb{Q}$ in $\mathbb{R}$
  \item Existence of square roots (completeness applied)
\end{enumerate}

\begin{remark*}[Primary sources]
Abbott, \textit{Understanding Analysis}, Chapter~1;
Rudin, \textit{Principles of Mathematical Analysis}, Chapter~1;
Ross, \textit{Elementary Analysis}, Chapter~1.
\end{remark*}

% ---------------------------------------------------------
% Content
% ---------------------------------------------------------
\input{volume-iii/analysis/bounding/notes/notes-completeness}
\input{volume-iii/analysis/bounding/notes/notes-bounds-extremal-values}
\input{volume-iii/analysis/bounding/notes/notes-bounding-monotonicty}

