% =========================================================
% Convergence: Sequences, Series, Asymptotics — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ Bounding (Bounds \& Completeness)
$\;\to\;$ \textbf{Convergence (Sequences)}
$\;\to\;$ Distance (Metric Spaces)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The Bounding chapter established that $\mathbb{R}$ is closed under
its own limiting behavior: every nonempty set bounded above has a
supremum.
Convergence puts that guarantee into motion.
A sequence is an infinite experiment in approximation, and the
completeness of $\mathbb{R}$ is what ensures the experiment can
succeed.

\medskip
\noindent\textbf{What this chapter builds.}
Rigorous $\varepsilon$-$N$ convergence, the algebra of limits,
monotone convergence, Cauchy sequences, subsequences and
Bolzano--Weierstrass, limit superior and inferior, infinite series,
convergence tests, and power series.

\medskip
\noindent\textbf{Where this leads.}
Metric spaces generalise the distance structure of $\mathbb{R}$ to
arbitrary sets. Every theorem proved here has a metric-space
analogue. This chapter is the prototype for all of it.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Role
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Convergence — The Dynamics of Approximation}},
  fonttitle=\small\bfseries
]
So far our structure has been static: sets, bounds, suprema.
Convergence introduces motion.

A sequence is not merely a function from $\mathbb{N}$ into $\mathbb{R}$.
It is a controlled infinite experiment --- a device that allows us
to observe approximation unfold step by step.

The $\varepsilon$-$N$ definition is the logical mechanism that
governs this approximation.
It trains us to manage nested quantifiers and to reason about
infinite tails.

Cauchy sequences deepen this picture.
They remove the target entirely: a sequence is Cauchy if its terms
eventually become arbitrarily close to \emph{one another}, with no
limit mentioned.
In $\mathbb{R}$, Cauchy convergence and ordinary convergence
coincide --- this equivalence expresses completeness in purely
metric language, bridging the order-based foundations of Bounding
with the distance-based abstraction of Distance.

The Bolzano--Weierstrass theorem closes the chapter's story.
Every bounded sequence has a convergent subsequence.
Completeness, originally a local supremum property, now manifests
globally: bounded infinite behavior cannot disperse indefinitely
--- it must accumulate.
This is the first hint of compactness.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Sequences $\longrightarrow$ Convergence
$\longrightarrow$ Subsequences $\longrightarrow$ Series
$\longrightarrow$ Asymptotics $\longrightarrow$ Applications}
\end{center}

\begin{enumerate}
  \item Sequences and boundedness
  \item Convergence and the algebra of limits
  \item Monotone sequences and the Monotone Convergence Theorem
  \item Cauchy sequences and completeness revisited
  \item Subsequences, Bolzano--Weierstrass, and the subsequence toolkit
  \item Recurrence inequalities and iterative control
  \item Limit superior and limit inferior
  \item Growth and asymptotics
  \item Infinite series and partial sums
  \item Absolute and conditional convergence
  \item Convergence tests (comparison, ratio, root, alternating)
  \item Rearrangements and power series
  \item Applications and worked examples
\end{enumerate}

\begin{remark*}[Primary sources]
Abbott, \textit{Understanding Analysis};
Rudin, \textit{Principles of Mathematical Analysis};
Pons, \textit{Real Analysis for the Undergraduate};
Ross, \textit{Elementary Analysis}.
\end{remark*}

% ---------------------------------------------------------
% Content
% ---------------------------------------------------------
\input{volume-iii/analysis/sequences/notes/sequences/index}
\input{volume-iii/analysis/sequences/notes/asymptotics/index}
\input{volume-iii/analysis/sequences/notes/series/index}
\input{volume-iii/analysis/sequences/notes/applications/index}
