\subsection{Bjørndahl}

\noindent
\textbf{Source.}
Adam Bjørndahl,
\textit{An Introduction to Classical and Modal Logics: The Outlines of Knowledge},
Chapter 1 (selected exercises).

\vspace{0.5em}

\begin{longtable}{|p{4.5cm}|p{8.5cm}|}
\hline
\textbf{Problem ID} & \textbf{Exercise (descriptor)} \\
\hline
\endfirsthead
\hline
\textbf{Problem ID} & \textbf{Exercise (descriptor)} \\
\hline
\endhead
\hline
\endfoot
\hline
\endlastfoot

% ------------------------------------------------------------
\hypertarget{PL-BJO-C01-E02a}{%
\hyperlink{proof-PL-BJO-C01-E02a}{\texttt{PL-BJO-C01-E02a}}}
&
Exercise 2(a): Recursive definition of $\mathrm{conn}(\varphi)$.
\\ \hline

% ------------------------------------------------------------
\hypertarget{PL-BJO-C01-RD01}{%
\hyperlink{proof-PL-BJO-C01-RD01}{\texttt{PL-BJO-C01-RD01}}}
&
Recursive definition: $\mathrm{atoms}(\varphi)$ = number of atomic formulas occurring in $\varphi$ (counting duplicates).
\\ \hline

\hypertarget{PL-BJO-C01-RD01}{%
\hyperlink{proof-PL-BJO-C01-RD01}{\texttt{PL-BJO-C01-RD01}}}
&
Recursive definition problem:
Recursively define a function $\mathrm{atoms}(\varphi)$ that returns the number
of atomic formulas occurring in $\varphi$ (counting duplicates).
\\ \hline

\hypertarget{PL-BJO-C01-RD02}{%
\hyperlink{proof-PL-BJO-C01-RD02}{\texttt{PL-BJO-C01-RD02}}}
&
Recursive definition problem:
Recursively define a function $\mathrm{neg}(\varphi)$ that returns the number
of occurrences of the negation connective $\neg$ in $\varphi$.
\\ \hline

\hypertarget{PL-BJO-C01-RD03}{%
\hyperlink{proof-PL-BJO-C01-RD03}{\texttt{PL-BJO-C01-RD03}}}
&
Recursive definition problem:
Recursively define a function $\mathrm{bin}(\varphi)$ that returns the number
of binary connectives occurring in $\varphi$
(i.e.\ occurrences of $\wedge$, $\vee$, and $\to$).
\\ \hline







\hypertarget{PL-BJO-C01-E02b}{%
\hyperlink{proof-PL-BJO-C01-E02b}{\texttt{PL-BJO-C01-E02b}}}
&
Exercise 2(b): Recursive definition of $\mathrm{sub}(\varphi)$ (subformulas).
\\ \hline

\hypertarget{PL-BJO-C01-E03}{%
\hyperlink{proof-PL-BJO-C01-E03}{\texttt{PL-BJO-C01-E03}}}
&
Exercise 3: Is $\mathrm{length}(\varphi)>\mathrm{rank}(\varphi)$? (induction or counterexample).
\\ \hline

\hypertarget{PL-BJO-C01-E04}{%
\hyperlink{proof-PL-BJO-C01-E04}{\texttt{PL-BJO-C01-E04}}}
&
Exercise 4: Is $\mathrm{rank}(\varphi)\ge\mathrm{conn}(\varphi)$? (induction or counterexample).
\\ \hline

\hypertarget{PL-BJO-C01-E05b}{%
\hyperlink{proof-PL-BJO-C01-E05b}{\texttt{PL-BJO-C01-E05b}}}
&
Exercise 5(b): Prove $\forall\varphi,\ f(\varphi) > g(\varphi)$.
\\ \hline

\hypertarget{PL-BJO-C01-E06}{%
\hyperlink{proof-PL-BJO-C01-E06}{\texttt{PL-BJO-C01-E06}}}
&
Exercise 6: A formula with $n$ connectives has at most $2n+1$ subformulas.
\\ \hline

\hypertarget{PL-BJO-C01-E07}{%
\hyperlink{proof-PL-BJO-C01-E07}{\texttt{PL-BJO-C01-E07}}}
&
Exercise 7: Prove $|\mathrm{sub}(\varphi)| \le 2^{\mathrm{rank}(\varphi)+1}-1$.
\\ \hline

\hypertarget{PL-BJO-C01-E10}{%
\hyperlink{proof-PL-BJO-C01-E10}{\texttt{PL-BJO-C01-E10}}}
&
Exercise 10: Show each listed formula scheme is a tautology.
\\ \hline

\hypertarget{PL-BJO-C01-E11}{%
\hyperlink{proof-PL-BJO-C01-E11}{\texttt{PL-BJO-C01-E11}}}
&
Exercise 11: Prove logical equivalence $\approx$ is an equivalence relation.
\\ \hline

\hypertarget{PL-BJO-C01-E12}{%
\hyperlink{proof-PL-BJO-C01-E12}{\texttt{PL-BJO-C01-E12}}}
&
Exercise 12: Substitution and the Substitution Theorem (propositional logic).
\\ \hline

\end{longtable}
