% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{R}$, Real Analysis
$\;\to\;$ Algebraic Structures
$\;\to\;$ Metric Spaces
$\;\to\;$ \textbf{Point-Set Topology}
$\;\to\;$ Measure Theory
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Metric spaces showed us that convergence and continuity can be defined
without coordinates --- only distance matters.
Point-set topology takes the final abstraction step: it discards
distance altogether, retaining only the family of ``open sets''
as the primitive notion.

\medskip
\noindent\textbf{What this chapter builds.}
We define topological spaces axiomatically, study the open/closed set
framework, and develop continuity, compactness, and connectedness
in full generality.
The Hausdorff, second-countable, and metric conditions are studied
as special cases that recover familiar properties.

\medskip
\noindent\textbf{Where this leads.}
Measure theory requires a $\sigma$-algebra of measurable sets, which
is a cousin of the topology.
Differential geometry and algebraic topology build on the topological
foundations established here.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Topology — Structural Roadmap
% File: 07-topology/topology.tex
% =========================================================

\subsection{Introduction to Topology}

% =========================================================
% Structural Roadmap
% =========================================================
\subsubsection*{Structural Roadmap}

The development of topology follows the same
definition--theorem--structure architecture used throughout
this project, but now abstracts the notion of
\emph{neighborhood structure} independently of distance.

Each major topic is organized as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems $\longrightarrow$ Consequences and Logical Structure}
\end{center}

The global progression is:

\begin{enumerate}
  \item Definition of a topological space
  \item Open and closed sets
  \item Interior, closure, and boundary
  \item Accumulation (limit) points
  \item Subspace topology
  \item Compactness via open covers
  \item Sequential vs. open-cover compactness
\end{enumerate}

\vspace{1em}

\begin{remark}
Topology isolates the concept of
\emph{open sets} and neighborhood structure
without reference to algebra, order, or distance.
\end{remark}

\begin{remark}
Many notions previously defined using sequences
(convergence, compactness, limit points)
are reinterpreted here in terms of open sets.
\end{remark}

% =========================================================
% I. Topological Structure
% =========================================================

\input{volume-iii/analysis/point-set-topology/notes/notes-topology}

% =========================================================
% Future Sections (To Be Developed)
% =========================================================

% Uncomment as content expands:

% \input{07-topology/notes-closure-interior}
% \input{07-topology/notes-compactness}
% \input{07-topology/notes-subspace}
% \input{07-topology/notes-product-topology}

% =========================================================
% Logical Positioning
% =========================================================

\begin{remark}[Structural Position]
Topology is introduced after metric spaces
to separate the idea of neighborhood structure
from the notion of distance.

This abstraction allows compactness,
continuity, and convergence to be studied
in their most general setting.
\end{remark}
