% =========================================================
% Distance: Metric Spaces — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ Real Analysis
$\;\to\;$ \textbf{Distance (Metric Spaces)}
$\;\to\;$ Nearness (Topology)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Real analysis developed convergence, Cauchy sequences, and completeness
specifically for $\mathbb{R}$, using $|x - y|$ as the distance.
Metric spaces ask: which of those theorems depend only on the
\emph{properties} of distance, not on the specific structure of $\mathbb{R}$?

\medskip
\noindent\textbf{What this chapter builds.}
We define metric spaces axiomatically and port the key analytic
concepts into this abstract setting.
The triangle inequality is the engine driving every result.

\medskip
\noindent\textbf{Where this leads.}
Topology abstracts further by replacing the metric with a collection
of open sets. Functional analysis studies metric spaces of functions.
Every normed vector space is a metric space.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Role
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Distance — Abstraction of Proximity}},
  fonttitle=\small\bfseries
]
Up to this point, order and absolute value have guided our intuition.
Distance was defined as~$|x - y|$, rooted in the order structure of~$\mathbb{R}$.

Metric spaces abstract this idea.
They retain only the notion of distance,
encoded through the triangle inequality,
and discard order, coordinates, and algebraic structure entirely.

The logical progression is:
\[
  \text{Absolute Value}
  \;\Rightarrow\;
  \text{Triangle Inequality}
  \;\Rightarrow\;
  \text{Metric}
  \;\Rightarrow\;
  \text{Open Balls}
  \;\Rightarrow\;
  \text{Convergence}.
\]

Convergence, Cauchy behavior, and completeness
can now be studied in settings far beyond~$\mathbb{R}$:
spaces of functions, sequences, matrices, and more.

Metric spaces are the first layer of structural generalization.
Everything proved here will be inherited by topology,
functional analysis, and every normed vector space.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Logical Structure}
\end{center}

The development of metric space theory follows the same
definition--theorem--structure architecture used in the real line
foundations, but now in a fully abstract setting.
The global progression is:

\begin{enumerate}
  \item Definition of a metric space and the triangle inequality
  \item Open balls and the metric topology
  \item Convergence in metric spaces
  \item Cauchy sequences and completeness
  \item Compactness and sequential compactness
  \item The real line as a canonical complete metric space
\end{enumerate}

\begin{remark*}[Primary sources — ordered progression]

\srccite{AbbottUnderstandingAnalysis}, 
\textit{Understanding Analysis};

\srccite{FondaModernIntroductionAnalysis}, 
\textit{A Modern Introduction to Mathematical Analysis};

\srccite{RossElementaryAnalysis}, 
\textit{Elementary Analysis};

\srccite{LeblBasicAnalysisI}, 
\textit{Basic Analysis I};

\srccite{PonsRealAnalysis}, 
\textit{Real Analysis for the Undergraduate};

\srccite{KolmogorovFominIntroAnalysis}, 
\textit{Introductory Real Analysis};

\srccite{BurkillFirstCourseAnalysis}, 
\textit{A First Course in Mathematical Analysis};

\srccite{TaoAnalysis1}, 
\textit{Analysis I};

\srccite{PughRealMathematicalAnalysis2}, 
\textit{Real Mathematical Analysis};

\srccite{BrucknerRealAnalysis}, 
\textit{Real Analysis};

\srccite{RudinPMA}, 
\textit{Principles of Mathematical Analysis};

\srccite{LeblBasicAnalysisII}, 
\textit{Basic Analysis II}.

\end{remark*}


% ---------------------------------------------------------
% Content
% ---------------------------------------------------------
% --- Metric space foundations ---
\input{volume-iii/analysis/metric-spaces/notes/metric-spaces/notes-metric-space}
\input{volume-iii/analysis/metric-spaces/notes/metric-spaces/notes-axiom-verification}
\input{volume-iii/analysis/metric-spaces/notes/metric-spaces/notes-properties-metric-space}
\input{volume-iii/analysis/metric-spaces/notes/metric-spaces/notes-multiple-metrics}
\input{volume-iii/analysis/metric-spaces/notes/metric-spaces/notes-unit-ball-characterisations}
\input{volume-iii/analysis/metric-spaces/notes/metrics/notes-properties-of-metrics}
\input{volume-iii/analysis/metric-spaces/notes/metrics/notes-completeness-not-topological}
\input{volume-iii/analysis/metric-spaces/notes/isometries/notes-isometries}
\input{volume-iii/analysis/metric-spaces/notes/metric-spaces/notes-real-analysis-bridge}

% --- Balls: the basic open neighborhoods ---
\input{volume-iii/analysis/metric-spaces/notes/balls/notes-balls}
\input{volume-iii/analysis/metric-spaces/notes/balls/notes-working-with-balls}
\input{volume-iii/analysis/metric-spaces/notes/balls/notes-closure-vs-closed-ball}

% --- Limit points (depend only on balls/metric) ---
\input{volume-iii/analysis/metric-spaces/notes/limit-points/notes-limit-points}
\input{volume-iii/analysis/metric-spaces/notes/limit-points/notes-derived-set-closure}

% --- Open/closed sets (use limit-point characterization) ---
\input{volume-iii/analysis/metric-spaces/notes/open-closed-sets/open-closed-sets}

% --- Convergence and the sequential/topological bridge ---
\input{volume-iii/analysis/metric-spaces/notes/convergence/notes-convergence}
\input{volume-iii/analysis/metric-spaces/notes/cauchy/notes-cauchy}
\input{volume-iii/analysis/metric-spaces/notes/completeness/notes-completeness}
\input{volume-iii/analysis/metric-spaces/notes/completeness/notes-co-cauchy}
\input{volume-iii/analysis/metric-spaces/notes/convergence/notes-sequential-characterizations}


\input{volume-iii/analysis/metric-spaces/notes/example-spaces/notes-example-spaces}

% Future sections (uncomment as content is written):

% \input{volume-iii/analysis/metric-spaces/notes/notes-cauchy}
% \input{volume-iii/analysis/metric-spaces/notes/notes-completeness}
% \input{volume-iii/analysis/metric-spaces/notes/notes-compactness}
