% =========================================================
% Vector Space Definition
% =========================================================

\section{Basic Definitions}

\begin{definition}[Vector Space --- informal description]
\label{def:vector-space-informal}
A vector space is a set equipped with addition and scalar multiplication
that behave like familiar vector arithmetic.
\end{definition}

\begin{remark*}[Fully quantified form]
\[ \text{A set closed under scalar multiplication and vector addition.} \]
\end{remark*}

\begin{remark*}[Flash]
\FlashQ{Describe a vector space informally.}
\FlashA{A set of objects (vectors) closed under addition and scalar multiplication over a field.}
\end{remark*}

\begin{remark*}
The informal description suggests structure, but precision requires axioms.
\end{remark*}

\begin{definition}[Vector Space]
\label{def:vector-space}
Let $\mathbb{F}$ be a field.
A \textbf{vector space} over $\mathbb{F}$ is a triple $(V, +, \cdot)$ where $V$ is a set,
$+: V \times V \to V$ is a binary operation called \textbf{vector addition}, and
$\cdot : \mathbb{F} \times V \to V$ is a binary operation called \textbf{scalar multiplication},
such that:

\begin{itemize}
    \item $(V, +)$ is an abelian group,
    \item $\alpha (u + v) = \alpha u + \alpha v$
          for all $\alpha \in \mathbb{F}$ and $u, v \in V$,
    \item $(\alpha + \beta) v = \alpha v + \beta v$
          for all $\alpha, \beta \in \mathbb{F}$ and $v \in V$,
    \item $(\alpha \beta) v = \alpha (\beta v)$
          for all $\alpha, \beta \in \mathbb{F}$ and $v \in V$,
    \item $1_{\mathbb{F}} v = v$ for all $v \in V$.
\end{itemize}

The elements of $V$ are called \textbf{vectors} and the elements of $\mathbb{F}$
are called \textbf{scalars}.
\end{definition}

\begin{remark*}[Fully quantified form]
\[ V\text{ vector space over }\mathbb{F} \iff \text{8 axioms: 4 additive, 4 scalar mult}. \]
\end{remark*}

\begin{remark*}[Flash]
\FlashQ{State the definition of a vector space.}
\FlashA{A set $V$ over field $\mathbb{F}$ with $+$ and scalar mult satisfying the 8 vector space axioms.}
\end{remark*}

\begin{remark*}[Expanded Axioms]
Expanding the abelian group structure of $(V,+)$ yields the full list of vector space axioms:

\begin{itemize}
\item \textbf{Associativity of addition:}
      $(u + v) + w = u + (v + w)$ for all $u,v,w \in V$.
\item \textbf{Commutativity of addition:}
      $u + v = v + u$ for all $u,v \in V$.
\item \textbf{Additive identity:}
      There exists $\mathbf{0} \in V$ such that $v + \mathbf{0} = v$ for all $v \in V$.
\item \textbf{Additive inverse:}
      For each $v \in V$ there exists $-v \in V$ such that $v + (-v) = \mathbf{0}$.
\item \textbf{Distributivity over vector addition:}
      $\alpha(u + v) = \alpha u + \alpha v$
      for all $\alpha \in \mathbb{F}$, $u,v \in V$.
\item \textbf{Distributivity over scalar addition:}
      $(\alpha + \beta)v = \alpha v + \beta v$
      for all $\alpha,\beta \in \mathbb{F}$, $v \in V$.
\item \textbf{Compatibility:}
      $(\alpha\beta)v = \alpha(\beta v)$
      for all $\alpha,\beta \in \mathbb{F}$, $v \in V$.
\item \textbf{Unit law:} $1_{\mathbb{F}} v = v$ for all $v \in V$.
\end{itemize}

Thus the structural definition compresses eight axioms into one abelian group condition
plus four scalar multiplication laws.
\end{remark*}

\begin{remark*}[Structural Summary]
A vector space combines an abelian group with a compatible scalar action:

\begin{center}
\begin{tabular}{l|c|c}
Structure & Additive Group & Scalar Action \\
\hline
Vector Space & Abelian & $\mathbb{F}$-linear
\end{tabular}
\end{center}

The scalar multiplication axioms express compatibility between the field $\mathbb{F}$
and the group $(V,+)$.
\end{remark*}

\begin{definition}[Zero vector]
\label{def:zero-vector}
Let $V$ be a vector space over a field $\mathbb{F}$.
The \emph{zero vector} of $V$, denoted $\mathbf{0}$, is the additive identity
of the abelian group $(V,+)$; that is, the unique element $\mathbf{0} \in V$
such that
\[
v + \mathbf{0} = v
\quad \text{for all } v \in V.
\]
\end{definition}

\begin{remark*}[Fully quantified form]
\[ \exists!\, \mathbf{0}\in V: \forall v\in V,\; v+\mathbf{0}=v. \]
\end{remark*}

\begin{remark*}[Flash]
\FlashQ{Define the zero vector.}
\FlashA{The unique element $\mathbf{0}\in V$ with $v+\mathbf{0}=v$ for all $v\in V$.}
\end{remark*}