% =========================================================
% Nearness: Point-Set Topology — Notes Index
% =========================================================

% ---------------------------------------------------------
% Breadcrumb
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{R}$, Real Analysis
$\;\to\;$ Metric Spaces
$\;\to\;$ \textbf{Nearness (Point-Set Topology)}
$\;\to\;$ Finiteness (Compactness)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Metric spaces showed us that convergence and continuity can be defined
without coordinates --- only distance matters.
Point-set topology takes the final abstraction step: it discards
distance altogether, retaining only the family of open sets
as the primitive notion.

\medskip
\noindent\textbf{What this chapter builds.}
We define the open-set framework, develop closure, interior, limit
points, and compactness, and establish the Heine--Borel theorem
for $\mathbb{R}^n$.

\medskip
\noindent\textbf{Where this leads.}
Compactness is the global finiteness principle that connects topology
back to analysis: continuous functions on compact sets attain their
extrema and are uniformly continuous, unlocking the Riemann integral.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Role
% ---------------------------------------------------------
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Role: Nearness — Neighborhood Structure Without Distance}},
  fonttitle=\small\bfseries
]
Metric spaces taught us that distance controls convergence.
Nearness --- point-set topology --- goes one step further.

It discards the metric entirely
and retains only the family of open sets:
the neighborhoods that determine local behavior.

Convergence, continuity, compactness, and limit points
can all be defined using only this neighborhood structure.

Topology reveals that many sequence-based results
were secretly statements about open sets.
The abstraction completes a three-stage descent:

\begin{itemize}
  \item Order $\;\to\;$ distance,
  \item Distance $\;\to\;$ neighborhoods,
  \item Neighborhoods $\;\to\;$ open sets.
\end{itemize}

Each stage removes a layer of extra structure
while preserving the phenomena that matter for analysis.
Topology isolates the minimal data needed to discuss
convergence and continuity.
It is the language in which the deepest theorems of analysis are stated,
even when their proofs appear to use $\varepsilon$--$\delta$.
\end{tcolorbox}
\vspace{1em}

% ---------------------------------------------------------
% Structural Roadmap
% ---------------------------------------------------------
\subsection*{Structural Roadmap}

\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Logical Structure}
\end{center}

The development of topology follows the same architecture used
throughout this project, but now abstracts neighborhood structure
independently of distance.
The global progression is:

\begin{enumerate}
  \item Open sets, closed sets, and their algebra
  \item Interior, closure, and boundary
  \item Limit points and the sequential characterization of closure
  \item Compactness via open covers
  \item Sequential compactness and the Heine--Borel theorem
  \item Connectedness
\end{enumerate}

\begin{remark*}[Primary sources]
Abbott, \textit{Understanding Analysis}, Chapter~3;
Pugh, \textit{Real Mathematical Analysis}, Chapter~2 §1--6;
Rudin, \textit{Principles of Mathematical Analysis}, Chapter~2.
\end{remark*}

% ---------------------------------------------------------
% Content
% ---------------------------------------------------------
\input{volume-iii/analysis/point-set-topology/notes/notes-topology}

% Future sections (uncomment as content is written):
% \input{volume-iii/analysis/point-set-topology/notes/notes-closure-interior}
% \input{volume-iii/analysis/point-set-topology/notes/notes-limit-points}
% \input{volume-iii/analysis/point-set-topology/notes/notes-compactness}
% \input{volume-iii/analysis/point-set-topology/notes/notes-connectedness}
