% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$
$\;\to\;$ \textbf{Real Numbers ($\mathbb{R}$)}
$\;\to\;$ Real Analysis
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The rational numbers gave us a dense ordered field, but they contain gaps:
there is no $q \in \mathbb{Q}$ with $q^2 = 2$.
The real number system fills these gaps. Two rigorous constructions achieve
this — Dedekind cuts and Cauchy sequences — and both yield the same unique
complete ordered field.

\medskip
\noindent\textbf{What this chapter builds.}
We lay down the axioms that characterise $\mathbb{R}$ as a complete ordered
field, develop the theory of bounds and suprema, and establish completeness
as the defining analytic property.
We then use completeness to build the full theory of sequences, series,
and their limiting behaviour.

\medskip
\noindent\textbf{Where this leads.}
Real analysis in Volume~III takes $\mathbb{R}$ as given and studies functions,
continuity, differentiation, and integration.
The field and order axioms established here propagate unchanged into
metric spaces, normed spaces, and abstract algebra.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\subsection*{Structural Roadmap}

The development follows the definition--theorem--structure architecture
used throughout this project.
Each major topic is organised as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Logical Structure}
\end{center}

The global progression is:
\begin{enumerate}
  \item Field and order axioms
  \item Intervals and convexity
  \item Bounds, suprema, and infima
  \item Completeness of $\mathbb{R}$ and its consequences
  \item Constructions of $\mathbb{R}$: Dedekind cuts and Cauchy sequences
  \item Sequences: definitions and boundedness
  \item Convergence and the algebra of limits
  \item Monotone sequences
  \item Cauchy sequences and completeness revisited
  \item Subsequences and Bolzano--Weierstrass
  \item Subsequence toolkit
  \item Recurrence inequalities and iterative control
  \item Limit superior and limit inferior
  \item Growth and asymptotics
  \item Infinite series
  \item Absolute and conditional convergence
  \item Series tests
  \item Rearrangements and power series
  \item Applications
\end{enumerate}

\vspace{1em}

\begin{remark}[Primary sources]
The axiomatic development follows Abbott, \textit{Understanding Analysis}
and Ross, \textit{Elementary Analysis}.
The Cauchy sequence construction of $\mathbb{R}$ follows Tao,
\textit{Analysis~I}, Chapters~5--6.
The Dedekind cut construction follows Rudin,
\textit{Principles of Mathematical Analysis}, Appendix to Chapter~1.
\end{remark}

\begin{remark}[Separation of concerns]
The foundations subfolder (axioms, bounds, completeness) develops the
\emph{algebraic and order structure} of $\mathbb{R}$.
The constructions subfolder explains \emph{how} $\mathbb{R}$ is built
from $\mathbb{Q}$, justifying that such an object exists and is unique.
The sequences and series subfolders develop \emph{analysis on} $\mathbb{R}$,
taking completeness as the key tool.
\end{remark}

% =========================================================
% Content
% =========================================================
\input{volume-ii/reals/notes/index}
