% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ $\mathbb{R}$
$\;\to\;$ Algebraic Structures
$\;\to\;$ \textbf{Linear Algebra}
$\;\to\;$ Topology
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Algebraic structures gave us the hierarchy of groups, rings, and fields.
A field is the scalar system: a set of numbers with addition and
multiplication. An abelian group is the additive backbone.
Linear algebra asks: what happens when you combine an abelian group
of ``vectors'' with scalar multiplication from a field?

\medskip
\noindent\textbf{What this chapter builds.}
We develop vector spaces from the axioms, then build the theory of
subspaces, linear combinations, span, linear independence, bases,
and dimension. Linear maps are the central objects of study ---
not matrices, which are only representations relative to a basis.

\medskip
\noindent\textbf{Where this leads.}
Every proof about eigenvalues, the spectral theorem, and inner product
spaces relies directly on the axioms established here.
Functional analysis generalises vector spaces to infinite dimensions;
topology provides the analytic tools for that generalisation.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Linear Algebra — Structural Roadmap
% Driver Text: Sheldon Axler, Linear Algebra Done Right
% =========================================================

\subsection{Linear Algebra}

% =========================================================
% Structural Roadmap
% =========================================================


\subsubsection*{Structural Roadmap}

The development of linear algebra in this project follows
the definition--theorem--structure architecture used
throughout the analysis volumes.

The primary driver is \textit{Linear Algebra Done Right}
by Sheldon Axler. The emphasis is on linear maps,
invariant structure, and conceptual clarity
rather than computational techniques.

Each major topic is organized as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems $\longrightarrow$ Consequences and Structural Insight}
\end{center}

The global progression is:

\begin{enumerate}
  \item Vector spaces and subspaces
  \item Linear combinations, span, and linear independence
  \item Bases and dimension
  \item Linear maps
  \item Null space and range
  \item Matrix representations of linear maps
  \item Eigenvalues and eigenvectors
  \item Invariant subspaces
  \item Triangularization and diagonalization
  \item Inner product spaces
  \item Orthogonality and projections
  \item Self-adjoint and normal operators
  \item Spectral theorem (finite-dimensional case)
\end{enumerate}

\vspace{1em}

\begin{remark}
This treatment prioritizes structural understanding over
algorithmic manipulation. Determinants are introduced only
after linear maps and eigenvalues have been conceptually
understood.
\end{remark}

\begin{remark}
The central object of study is not matrices,
but linear maps between vector spaces.
Matrices serve only as representations relative to a basis.
\end{remark}

\begin{remark}[Structural Position]
Linear algebra is developed independently of analysis.
Its structural viewpoint will later support operator theory,
Hilbert spaces, and functional analysis.
\end{remark}

% =========================================================
% I. Core Linear Structure
% =========================================================

\input{volume-iii/algebra/linear-algebra/notes/vector-spaces/vector-spaces}

% =========================================================
% II. Linear Maps
% =========================================================

% \input{volume-iii/algebra/linear-algebra/notes/notes-linear-independence}
% \input{volume-iii/algebra/linear-algebra/notes/notes-basis-dimension}
% \input{volume-iii/algebra/linear-algebra/notes/notes-linear-maps}
% \input{volume-iii/algebra/linear-algebra/notes/notes-null-range}
% \input{volume-iii/algebra/linear-algebra/notes/notes-matrix-representations}

% =========================================================
% III. Eigenstructure
% =========================================================

% \input{volume-iii/algebra/linear-algebra/notes/notes-eigenvalues}
% \input{volume-iii/algebra/linear-algebra/notes/notes-invariant-subspaces}
% \input{volume-iii/algebra/linear-algebra/notes/notes-diagonalization}

% =========================================================
% IV. Inner Product Structure
% =========================================================

% \input{volume-iii/algebra/linear-algebra/notes/notes-inner-products}
% \input{volume-iii/algebra/linear-algebra/notes/notes-orthogonality}
% \input{volume-iii/algebra/linear-algebra/notes/notes-spectral-theorem}

% =========================================================
% Future Sections (To Be Developed)
% =========================================================

% \input{volume-iii/algebra/linear-algebra/notes/notes-determinants}
% \input{volume-iii/algebra/linear-algebra/notes/notes-jordan-form}
% \input{volume-iii/algebra/linear-algebra/notes/notes-advanced-operators}

% =========================================================
% Logical Positioning
% =========================================================

