% =========================================================
% Rational Approximation
% =========================================================

\section{Rational Approximation}
\label{sec:rational-approximation}

\subsection{Definitions and Theorems}

One of the most important structural facts about the rational numbers is that they can approximate
other numbers arbitrarily well.  This is the point at which the density of $\mathbb{Q}$ begins to turn
into analysis.

\begin{theorem}[Rational approximation]
\label{thm:rational-approximation}
For every real number $x$ and every $\varepsilon>0$, there exists $r\in\mathbb{Q}$ such that
\[
|x-r|<\varepsilon.
\]
\end{theorem}

\begin{proof}
Apply the density theorem to the interval $(x-\varepsilon,x+\varepsilon)$.  The rational number found
in that interval satisfies the desired inequality.
\end{proof}

\begin{remark}
This theorem says that rational numbers are dense not only in the order-theoretic sense of lying
between real numbers, but also in the metric sense of lying arbitrarily close to any prescribed real
number.
\end{remark}

\begin{remark}[Constructive approximation methods]
There are several useful ways to approximate a real number by rational numbers.
\begin{enumerate}
  \item \textbf{Decimal truncation.}  If $x=1.4142135\dots$, then
  \[
  1.4,\quad 1.41,\quad 1.414,\quad 1.4142
  \]
  are rational approximations.
  \item \textbf{Mediants.}  If
  \[
  \frac{a}{b}<x<\frac{c}{d},
  \]
  then the mediant
  \[
  \frac{a+c}{b+d}
  \]
  gives a new rational approximation between the two old ones.
  \item \textbf{Continued fractions.}  Continued fractions often produce exceptionally good rational
  approximations and are closely related to the Stern--Brocot tree.
\end{enumerate}
\end{remark}

\begin{remark}[Approximation by sequences]
A real number can often be approximated naturally by a sequence of rationals.  For example,
\[
1,\ 1.4,\ 1.41,\ 1.414,\ 1.4142,\dots
\]
is a sequence of rational approximations to $\sqrt{2}$.  This point of view leads directly to Cauchy
sequences and to the construction of the real numbers.
\end{remark}

\subsection{Dirichlet's Approximation Theorem}

Density alone says that rationals come arbitrarily close to a real number, but it says nothing about
\emph{how efficient} those approximations can be.  Dirichlet's theorem gives a surprisingly strong
quantitative answer.

\begin{theorem}[Dirichlet's Approximation Theorem]
\label{thm:dirichlet-approx}
For every real number $x$ there exist infinitely many rational numbers $p/q$ with $p\in\mathbb{Z}$ and
$q\in\mathbb{N}$ such that
\[
\left|x-\frac{p}{q}\right|<\frac{1}{q^2}.
\]
\end{theorem}

\begin{remark}[Interpretation]
If the denominator $q$ is moderately large, the error bound $1/q^2$ can already be very small.  In
this sense the rational numbers approximate real numbers far better than density by itself might
suggest.
\end{remark}

\begin{remark}[Idea of proof]
Consider the fractional parts of
\[
x,2x,3x,\dots,(N+1)x.
\]
These $N+1$ numbers lie in the interval $[0,1)$, which is divided into $N$ equal subintervals.
By the pigeonhole principle two of the fractional parts must land in the same subinterval.  Their
difference produces integers $p$ and $q$ for which
\[
\left|x-\frac{p}{q}\right|<\frac{1}{qN}.
\]
A suitable choice of $N$ then yields the stated estimate.
\end{remark}

\begin{remark}[Conceptual significance]
The study of how well real numbers can be approximated by rationals is called
\emph{Diophantine approximation}.  Dirichlet's theorem is one of its basic foundational results and
forms a natural bridge from elementary rational arithmetic to deeper number theory.
\end{remark}


\input{volume-ii/rationals/notes/figure-geometry-rational-approximation}

% =========================================================
% Exposition: The Geometry of Rational Approximation
% For inclusion in Volume III, Bounding chapter,
% following the Density of Q and Density of Irrationals.
% =========================================================

\section*{Reading the Diagram: A Walking Tour}

The figure above is a single picture with three layers, each telling
the same story from a different vantage point.
Take a moment to orient yourself before reading further.

\bigskip

\section*{Why rational approximation matters}

You have just proved that the rationals are dense in $\mathbb{R}$:
between any two real numbers lies a rational one.
That result tells you approximation is \emph{possible}, but it says
nothing about how \emph{efficient} those approximations can be made.

Here is the sharper question: given an irrational number $x$, how
close can a fraction $p/q$ get to $x$, and how does the quality of
the approximation depend on the size of the denominator $q$?
This is the question the diagram is designed to answer geometrically.

\bigskip

\section*{Layer 1 (top): The Farey gap identity}

The number line at the top shows two fractions in \emph{lowest terms},
$a/b$ and $c/d$, that are \emph{Farey neighbors}: that is, they satisfy
$|bc - ad| = 1$ (equivalently, $bc - ad = 1$ if $a/b < c/d$).
The blue tick marks the \emph{mediant} $(a+c)/(b+d)$, which lies
strictly between them.

\medskip

The key identity, shown by the double-headed arrow below the line, is
\[
\frac{c}{d} - \frac{a}{b} = \frac{1}{bd}.
\]
This is not an approximation --- it is \emph{exact}.
Farey neighbors are not merely close; their distance is completely
determined by their denominators.
A moment's algebra confirms it:
\[
\frac{c}{d} - \frac{a}{b}
= \frac{bc - ad}{bd}
= \frac{1}{bd},
\]
where the last step uses the Farey neighbor condition $bc - ad = 1$.

\medskip

The blue annotations above the line show what the mediant does to this
gap.  It does not split the interval in half; it splits it into two
pieces of \emph{unequal} size:
\[
\frac{a+c}{b+d} - \frac{a}{b} = \frac{1}{b(b+d)},
\qquad
\frac{c}{d} - \frac{a+c}{b+d} = \frac{1}{d(b+d)}.
\]
The larger denominator produces the smaller subgap.
This asymmetry is the engine of the whole theory: inserting mediants
refines intervals, but does so unevenly, always shrinking more on the
side with the larger existing denominator.

\medskip

\begin{remark*}[Check the algebra yourself]
Verify both subgap identities by direct computation.
You will use $bc - ad = 1$ in each case.
This is a one-line calculation and worth doing at least once.
\end{remark*}

\bigskip

\section*{Layer 2 (middle): The Stern--Brocot tree}

If you apply mediant insertion \emph{repeatedly}, starting from
$0/1$ and $1/0$ (a formal symbol for $\infty$, used to bookend the
positive rationals), you generate the \emph{Stern--Brocot tree}.

\medskip

The middle panel shows the first two levels of this construction.
\begin{itemize}
  \item \textbf{Level 0.} The seeds: $0/1$, $1/1$, $1/0$.
  \item \textbf{Level 1.} Insert mediants between adjacent seeds.
        Between $0/1$ and $1/1$ the mediant is
        $\frac{0+1}{1+1} = \frac{1}{2}$.
        Between $1/1$ and $1/0$ the mediant is
        $\frac{1+1}{1+0} = \frac{2}{1}$.
  \item \textbf{Level 2.} Insert mediants between every adjacent pair
        at level 1, yielding $1/3$, $2/3$, $3/2$, $3/1$.
\end{itemize}

\medskip

What makes the Stern--Brocot tree remarkable is the following theorem
(whose proof we defer): \emph{every positive rational number appears
in the tree exactly once, already in lowest terms.}
The tree is therefore a complete, duplicate-free enumeration of all
positive rationals, built out of nothing but repeated mediant
insertion.

\medskip

Notice how the fractions spread across the number line: small
denominators appear at the top (coarse level) and large denominators
appear further down (fine level).
The tree encodes the idea that simple fractions --- those with small
denominators --- are the best approximations to any given real number,
and complex fractions (large denominators) are needed only to
approximate with high precision.

\bigskip

\section*{Layer 3 (bottom): Dirichlet approximation}

The bottom number line makes the approximation viewpoint explicit.
The red tick is an irrational number $x$.
Around it float four rational approximants $p_1/q_1, p_2/q_2,
p_3/q_3, p_4/q_4$ with increasing denominators, each closer to $x$
than the last.

\medskip

The three concentric dashed circles are a visual metaphor for
the Dirichlet bound
\[
\left| x - \frac{p}{q} \right| < \frac{1}{q^2}.
\]
Each circle represents a shrinking window: as $q$ grows, the window
$1/q^2$ tightens around $x$, forcing the approximants to cluster.

\medskip

\textbf{Why $1/q^2$?}
The Farey gap identity tells us that two Farey neighbors at level $q$
are separated by $1/(q \cdot q')$ where $q' \geq q$.
The irrational $x$ must land in one of these subintervals, and the
nearer neighbor satisfies the $1/q^2$ bound.
So Dirichlet's theorem is not a magical accident; it is the Farey gap
identity applied to the interval containing $x$.

\medskip

\begin{remark*}[Dirichlet's theorem]
For any irrational $x$ and any $N \in \mathbb{N}$, there exists
$p/q$ with $1 \leq q \leq N$ such that
\[
\left| x - \frac{p}{q} \right| < \frac{1}{qN} \leq \frac{1}{q^2}.
\]
The proof uses the Pigeonhole Principle on the $N+1$ fractional parts
$\{kx\}$ for $k = 0, 1, \ldots, N$ distributed among $N$ subintervals
of length $1/N$.
\end{remark*}

\bigskip

\section*{The hidden unity}

The diagram's caption says these are ``different manifestations of the
same hidden geometry of the rational numbers.''
What is that geometry, precisely?

\medskip

The answer is the Farey identity $bc - ad = 1$.
\begin{itemize}
  \item In \textbf{Layer 1}, it gives the exact gap $1/(bd)$.
  \item In \textbf{Layer 2}, it is preserved by mediant insertion:
        if $a/b$ and $c/d$ are Farey neighbors, then $a/b$ and
        $(a+c)/(b+d)$ are Farey neighbors, and so is
        $(a+c)/(b+d)$ and $c/d$.
        The tree grows by propagating this relation.
  \item In \textbf{Layer 3}, it implies the Dirichlet bound, because
        $x$ lies in a Farey interval whose endpoints satisfy
        $bc - ad = 1$.
\end{itemize}

\medskip

This invariant --- an integer determinant equal to $1$ --- is the
thread connecting all three panels.
It is the same identity that appears in the theory of continued
fractions, in the modular group $\mathrm{SL}(2,\mathbb{Z})$, and in
the study of lattice points.
The diagram is your first glimpse of a deep structure that recurs
throughout number theory and geometry.

\bigskip

\section*{What to take away}

After studying this diagram, you should be able to answer the
following questions without looking anything up:

\begin{enumerate}
  \item Why is the gap between Farey neighbors $a/b$ and $c/d$
        exactly $1/(bd)$ rather than approximately $1/(bd)$?
  \item Given two Farey neighbors, which subgap --- left or right ---
        is smaller after mediant insertion, and why?
  \item Why does increasing the denominator $q$ make a better
        rational approximation to an irrational $x$?
  \item What is the connection between the Stern--Brocot tree and
        continued fractions?
        (You do not know the answer yet, but formulating the question
        precisely is the right first step.)
\end{enumerate}

\noindent
If you can answer (1)--(3) from the diagram alone, without algebra,
you have genuinely understood the picture.
Question (4) points to where we are going.










