% =========================================================
% Algebraic Structures — Structural Roadmap
% Driver Text: Joseph Gallian, Contemporary Abstract Algebra
% =========================================================
\subsection{Algebraic Structures}

% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Real Analysis
$\;\to\;$ \textbf{Algebraic Structures}
$\;\to\;$ Linear Algebra
$\;\to\;$ Topology
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Sets and functions gave us the language of mathematical structure,
and proof techniques gave us the tools to reason about it.
Real analysis showed what rigorous reasoning looks like on a concrete
number system. Algebraic structures now ask the deeper question:
what is the \emph{minimal} set of axioms needed for the most important
algebraic phenomena --- identity, inverses, and arithmetic --- to occur?

\medskip
\noindent\textbf{What this chapter builds.}
We construct the hierarchy of algebraic structures: groups, rings,
and fields. Each is defined by progressively richer axiom systems.
Fields are the scalars over which vector spaces are built,
and groups are the additive backbone of every vector space.

\medskip
\noindent\textbf{Where this leads.}
Linear algebra inherits the field and group axioms directly:
a vector space is an abelian group equipped with a scalar multiplication
by a field, and every proof about vector spaces draws on both simultaneously.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\subsubsection*{Structural Roadmap}

The development of algebraic structures in this project follows
the definition--theorem--structure architecture used
throughout the analysis volumes.

The primary driver is \textit{Contemporary Abstract Algebra}
by Joseph Gallian. The emphasis is on axiom systems,
structural consequences, and the hierarchy of algebraic objects
rather than computational techniques.

Each major topic is organized as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Structural Insight}
\end{center}

The global progression is:
\begin{enumerate}
  \item Groups and abelian groups
  \item Subgroups
  \item Rings
  \item Integral domains
  \item Fields and their properties
\end{enumerate}

\vspace{1em}

\begin{remark}[Structural Position]
The structures developed here are not studied for their own sake alone.
Groups provide the additive structure of every vector space.
Fields provide the scalars. The interaction between them
is the subject of linear algebra.
\end{remark}

\begin{remark}[Dependency Note]
The uniqueness theorems proved here --- uniqueness of identity,
uniqueness of inverses --- are the same theorems invoked in
vector space proofs. They are proved once here and cited by
theorem number thereafter.
\end{remark}

% =========================================================
% I. Groups
% =========================================================
\input{intro-algebraic-structures/notes/groups/index}
\clearpage
\section{Worksheets}
\input{intro-algebraic-structures/worksheets}

\section{Proofs}algebraic-structures
\input{intro-algebraic-structures/proofs/index}

\section{Capstone}
\input{intro-algebraic-structures/capstone}
