% =========================================================
% Limits Superior and Inferior of Functions — Notes
% Chapter: continuity
% Volume III · Analysis
% =========================================================

\subsection{Limits Superior and Inferior of Functions}
\label{sec:limsup-liminf-functions}

\begin{tcolorbox}[title={Toolkit --- Limits Superior and Inferior of Functions}]
\begin{itemize}
  \item $\limsup_{x\to x_0}f(x)=\inf_{\delta>0}
    \sup\{f(x):0<|x-x_0|<\delta,\,x\in E\}$.
  \item Both always exist in $[-\infty,+\infty]$.
  \item $\limsup\geq\liminf$ always.
  \item $\operatorname{FunctionLimit}(f,x_0,L)$ exists
    $\iff\limsup=\liminf=L\in\mathbb{R}$.
  \item Direct analogue of sequence $\limsup$/$\liminf$: replace tail
    $\{a_n:n\geq N\}$ with punctured neighbourhood
    $\{f(x):\operatorname{InputTolerance}(x,x_0,\delta)\}$.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title={Definition --- Limits Superior and Inferior}]
\begin{definition}[Limits Superior and Inferior of a Function]
\label{def:limsup-liminf-function}
Let $f:E\to\mathbb{R}$, $x_0$ an accumulation point of $E$.
\[
  \limsup_{x\to x_0}f(x)
  \;:=\;
  \inf_{\delta>0}
  \sup\bigl\{\,f(x):x\in E,\;\operatorname{InputTolerance}(x,x_0,\delta)\,\bigr\}.
\]
\[
  \liminf_{x\to x_0}f(x)
  \;:=\;
  \sup_{\delta>0}
  \inf\bigl\{\,f(x):x\in E,\;\operatorname{InputTolerance}(x,x_0,\delta)\,\bigr\}.
\]
\end{definition}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
  \limsup_{x\to x_0}f(x)
  = \inf_{\delta>0}
    \sup\{f(x):x\in E\cap V_\delta^*(x_0)\}.
\]
\end{remark*}

\begin{remark*}[Analogy with sequence $\limsup$/$\liminf$]
Sequence: $\limsup_{n\to\infty}a_n=\inf_N\sup_{n\geq N}a_n$.
Function: replace tail $\{a_n:n\geq N\}$ with punctured neighbourhood.
Structure identical; indexing set changes.
\end{remark*}

\begin{remark*}[Interpretation]
$\limsup$ is the largest value that $f$ comes arbitrarily close to
as $x\to x_0$. $\liminf$ is the infimum. If they coincide at a
finite value, the limit exists.
\end{remark*}

\begin{proposition}[$\limsup\geq\liminf$]
\label{prop:limsup-geq-liminf-function}
$\limsup_{x\to x_0}f(x)\geq\liminf_{x\to x_0}f(x)$.

\smallskip\noindent
\hyperref[prf:limsup-geq-liminf-function]{\textit{Go to proof.}}
\end{proposition}

\begin{proposition}[Limit Exists iff $\limsup=\liminf$]
\label{prop:limsup-liminf-limit-criterion}
$\operatorname{FunctionLimit}(f,x_0,L)\iff
\limsup_{x\to x_0}f(x)=\liminf_{x\to x_0}f(x)=L\in\mathbb{R}$.

\smallskip\noindent
\hyperref[prf:limsup-liminf-limit-criterion]{\textit{Go to proof.}}
\end{proposition}

\begin{remark*}[Standard quantified statement]
\[
  \operatorname{FunctionLimit}(f,x_0,L)
  \;\iff\;
  \limsup_{x\to x_0}f(x)=L=\liminf_{x\to x_0}f(x).
\]
\end{remark*}

\begin{remark*}[Interpretation]
Canonical failure: $f(x)=\sin(1/x)$ at $x_0=0$: $\limsup=1$,
$\liminf=-1$, no limit.
\end{remark*}
