% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ $\mathbb{R}$, Real Analysis
$\;\to\;$ Algebraic Structures
$\;\to\;$ \textbf{Metric Spaces}
$\;\to\;$ Topology
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Real analysis developed convergence, Cauchy sequences, and completeness
specifically for $\mathbb{R}$, using $|x - y|$ as the distance.
Metric spaces ask: which of those theorems depend only on
the \emph{properties} of distance, not on the specific structure of $\mathbb{R}$?

\medskip
\noindent\textbf{What this chapter builds.}
We define metric spaces axiomatically, then port the key analytic
concepts --- open and closed sets, sequences and limits, completeness,
compactness, and connectedness --- into this abstract setting.
The triangle inequality is the engine driving every result.

\medskip
\noindent\textbf{Where this leads.}
Topology abstracts further by replacing the metric with a collection
of open sets. Functional analysis studies metric spaces of functions.
Every normed vector space is a metric space.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Metric Spaces — Structural Roadmap
% File: 06-metric-spaces/metric-spaces.tex
% =========================================================

\subsection{Metric Spaces}

% =========================================================
% Structural Roadmap
% =========================================================
\subsubsection*{Structural Roadmap}

The development of metric space theory follows the same
definition--theorem--structure architecture used in the real line
foundations, but now in a fully abstract setting.

Each major topic is organized as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems $\longrightarrow$ Consequences and Logical Structure}
\end{center}

The global progression is:

\begin{enumerate}
  \item Definition of a metric space
  \item Open balls
  \item Convergence in metric spaces
  \item Cauchy sequences and completeness
  \item Compactness and sequential compactness
  \item The real line as a complete metric space
\end{enumerate}

\vspace{1em}

\begin{remark}
This section abstracts the distance structure previously used on
$\mathbb{R}$:
\[
d(x,y) = |x-y|.
\]
The real line now becomes a special case of a metric space.
\end{remark}

\begin{remark}
Metric space theory isolates the notion of distance
from order structure and supremum-based completeness.
It allows convergence and Cauchy behavior to be studied
independently of algebraic or order properties.
\end{remark}

% =========================================================
% I. Metric Structure
% =========================================================

\input{volume-iii/analysis/metric-spaces/notes/notes-metric-space}

% =========================================================
% Future Sections (To Be Developed)
% =========================================================

% Uncomment as content expands:

% \input{06-metric-spaces/notes-convergence-metric}
% \input{06-metric-spaces/notes-cauchy-metric}
% \input{06-metric-spaces/notes-completeness-metric}
% \input{06-metric-spaces/notes-compactness-metric}

% =========================================================
% Logical Positioning
% =========================================================

\begin{remark}[Structural Position]
Metric spaces are introduced only after sequences,
convergence, and completeness have been fully understood
in $\mathbb{R}$.

This ensures that abstraction follows mastery rather than
preceding it.
\end{remark}
