% =========================================================
% Vector Spaces
% =========================================================

% ---------------------------------------------------------
\subsubsection{Preliminary Definitions}
% ---------------------------------------------------------

\begin{definition}[Abelian group]
A set $G$ with a binary operation $\ast : G \times G \to G$
is an \emph{abelian group} if:

\begin{enumerate}
\item \textbf{Associativity:}
\[
(a \ast b) \ast c = a \ast (b \ast c)
\quad\text{for all } a,b,c \in G.
\]

\item \textbf{Identity element:}
There exists $e \in G$ such that
\[
a \ast e = e \ast a = a
\quad\text{for all } a \in G.
\]

\item \textbf{Inverse element:}
For each $a \in G$ there exists $b \in G$ such that
\[
a \ast b = b \ast a = e.
\]

\item \textbf{Commutativity:}
\[
a \ast b = b \ast a
\quad\text{for all } a,b \in G.
\]
\end{enumerate}
\end{definition}

\begin{remark}
An abelian group is simply a group whose operation is commutative.
\end{remark}

\begin{definition}[Field — structural form]
A \emph{field} is a set $\mathbb{F}$ equipped with two operations
$+$ and $\cdot$ such that:

\begin{itemize}
\item $(\mathbb{F}, +)$ is an abelian group,
\item $(\mathbb{F}\setminus\{0\}, \cdot)$ is an abelian group,
\item Multiplication distributes over addition.
\end{itemize}
\end{definition}

\begin{remark}[Expanded Axioms]
Expanding the abelian group axioms yields the familiar list:

\begin{itemize}
\item Associativity, commutativity, identity, and inverse for addition.
\item Associativity, commutativity, identity, and inverse (for nonzero elements) for multiplication.
\item The distributive law.
\end{itemize}

Thus the usual field axioms are not independent assumptions,
but consequences of the two abelian group structures together with distributivity.
\end{remark}

\begin{remark}[Structural Summary]
A field consists of two compatible abelian group structures:

\begin{center}
\begin{tabular}{l|c|c}
Structure & Additive Group & Multiplicative Structure \\
\hline
Field & Abelian & Abelian (on nonzero elements)
\end{tabular}
\end{center}

Distributivity links the two operations.
\end{remark}

% ---------------------------------------------------------
\subsubsection{The Complex Numbers}

\begin{definition}[Complex Numbers]
The set of \emph{complex numbers}, denoted $\mathbb{C}$, is defined as
\[
\mathbb{C} := \{(a,b) : a,b \in \mathbb{R}\},
\]
with operations defined by
\[
(a,b) + (c,d) := (a+c,\; b+d),
\]
\[
(a,b)\cdot(c,d) := (ac - bd,\; ad + bc).
\]
\end{definition}

\begin{remark}
We identify $(a,b)$ with the symbol
\[
a + bi,
\]
where $i := (0,1)$ satisfies
\[
i^2 = -1.
\]
\end{remark}

\begin{remark}
With these operations, $\mathbb{C}$ is a field containing $\mathbb{R}$
as the subset $\{(a,0): a \in \mathbb{R}\}$.
\end{remark}

\begin{remark}[Field structure of $\mathbb{C}$]
With the operations defined above, the set $\mathbb{C}$
satisfies the field axioms.

That is:
\begin{itemize}
\item $(\mathbb{C}, +)$ is an abelian group,
\item $(\mathbb{C}\setminus\{0\}, \cdot)$ is an abelian group,
\item multiplication distributes over addition.
\end{itemize}

Hence $\mathbb{C}$ is a field.
\end{remark}

% ---------------------------------------------------------
\subsubsection{The Space $\mathbb{C}^n$}

\begin{definition}[$\mathbb{C}^n$]
Let $n \in \mathbb{N}$.
The set
\[
\mathbb{C}^n := \{(z_1, \dots, z_n) : z_i \in \mathbb{C}\}
\]
is the set of ordered $n$-tuples of complex numbers.

Addition and scalar multiplication are defined componentwise:
\[
(z_1,\dots,z_n) + (w_1,\dots,w_n)
:=
(z_1+w_1,\dots,z_n+w_n),
\]
\[
\lambda (z_1,\dots,z_n)
:=
(\lambda z_1,\dots,\lambda z_n),
\quad
\lambda \in \mathbb{C}.
\]
\end{definition}

\begin{remark}
With these operations, $\mathbb{C}^n$ is a vector space over the field $\mathbb{C}$.
\end{remark}

% ---------------------------------------------------------
\subsubsection{Basic Definitions}

\begin{definition}[Vector Space — informal description]
A vector space is a set equipped with addition and scalar multiplication
that behave like familiar vector arithmetic.
\end{definition}

\begin{remark}
The informal description suggests structure, but precision requires axioms.
\end{remark}

\begin{definition}[Vector Space]
Let $\mathbb{F}$ be a field.
A \textbf{vector space} over $\mathbb{F}$ is a triple $(V, +, \cdot)$ where $V$ is a set,
$+: V \times V \to V$ is a binary operation called \textbf{vector addition}, and
$\cdot : \mathbb{F} \times V \to V$ is a binary operation called \textbf{scalar multiplication}, such that:
\begin{itemize}
    \item $(V, +)$ is an abelian group,
    \item $\alpha (u + v) = \alpha u + \alpha v$ for all $\alpha \in \mathbb{F}$ and $u, v \in V$,
    \item $(\alpha + \beta) v = \alpha v + \beta v$ for all $\alpha, \beta \in \mathbb{F}$ and $v \in V$,
    \item $(\alpha \beta) v = \alpha (\beta v)$ for all $\alpha, \beta \in \mathbb{F}$ and $v \in V$,
    \item $1_{\mathbb{F}} v = v$ for all $v \in V$.
\end{itemize}
The elements of $V$ are called \textbf{vectors} and the elements of $\mathbb{F}$ are called \textbf{scalars}.
\end{definition}

\begin{remark}[Expanded Axioms]
Expanding the abelian group structure of $(V,+)$ yields the full list of vector space axioms:

\begin{itemize}
\item \textbf{Associativity of addition:} $(u + v) + w = u + (v + w)$ for all $u,v,w \in V$.
\item \textbf{Commutativity of addition:} $u + v = v + u$ for all $u,v \in V$.
\item \textbf{Additive identity:} There exists $\mathbf{0} \in V$ such that $v + \mathbf{0} = v$ for all $v \in V$.
\item \textbf{Additive inverse:} For each $v \in V$ there exists $-v \in V$ such that $v + (-v) = \mathbf{0}$.
\item \textbf{Distributivity over vector addition:} $\alpha(u + v) = \alpha u + \alpha v$ for all $\alpha \in \mathbb{F}$, $u,v \in V$.
\item \textbf{Distributivity over scalar addition:} $(\alpha + \beta)v = \alpha v + \beta v$ for all $\alpha,\beta \in \mathbb{F}$, $v \in V$.
\item \textbf{Compatibility:} $(\alpha\beta)v = \alpha(\beta v)$ for all $\alpha,\beta \in \mathbb{F}$, $v \in V$.
\item \textbf{Unit law:} $1_{\mathbb{F}} v = v$ for all $v \in V$.
\end{itemize}

Thus the structural definition compresses eight axioms into one abelian group condition
plus four scalar multiplication laws.
\end{remark}

\begin{remark}[Structural Summary]
A vector space combines an abelian group with a compatible scalar action:

\begin{center}
\begin{tabular}{l|c|c}
Structure & Additive Group & Scalar Action \\
\hline
Vector Space & Abelian & $\mathbb{F}$-linear
\end{tabular}
\end{center}

The scalar multiplication axioms express compatibility between the field $\mathbb{F}$
and the group $(V,+)$.
\end{remark}

\begin{remark}
With the componentwise operations defined above, $\mathbb{F}^n$ is a vector space
over $\mathbb{F}$ for any field $\mathbb{F}$ and any $n \in \mathbb{N}$.
\end{remark}

% ---------------------------------------------------------
\subsubsection{Tuples and Lists}

\begin{definition}[$n$-tuple]
Let $n \in \mathbb{N}$ and let $X$ be a set.
An \emph{$n$-tuple} of elements of $X$ is a function
\[
f : \{1,2,\dots,n\} \to X.
\]
We write such a function as
\[
(x_1,\dots,x_n),
\]
where $x_i := f(i)$ for each $i$.
\end{definition}

\begin{definition}[Finite list]
Let $X$ be a set.
A \emph{finite list} of elements of $X$ is a function
\[
f : \{1,2,\dots,n\} \to X
\]
for some $n \in \mathbb{N}$.
\end{definition}

\begin{definition}[Set]
A \emph{set} is a collection of distinct elements
with no inherent ordering.
\end{definition}

\begin{example}
Consider the elements $1$ and $2$.

The set
\[
\{1,2,2\}
\]
is equal to
\[
\{1,2\},
\]
because sets do not record duplicates.

However, the list
\[
(1,2,2)
\]
is different from
\[
(1,2),
\]
because lists record both order and repetition.
\end{example}

\begin{definition}[Coordinate]
Let $\mathbb{F}$ be a field and let
\[
v = (v_1,\dots,v_n) \in \mathbb{F}^n.
\]
The element $v_i$ is called the \emph{$i$th coordinate} of $v$.
\end{definition}

% ---------------------------------------------------------
\subsubsection{Structural Interpretation of $\mathbb{F}^n$}

\begin{definition}[The space $\mathbb{F}^n$ as a function space]
Let $\mathbb{F}$ be a field and let $n \in \mathbb{N}$.
Define
\[
\mathbb{F}^n := \{ f : \{1,\dots,n\} \to \mathbb{F} \}.
\]
\end{definition}

\begin{remark}
An element $v \in \mathbb{F}^n$ is therefore a function
\[
v : \{1,\dots,n\} \to \mathbb{F}.
\]
We write
\[
v = (v_1,\dots,v_n),
\quad \text{where } v_i := v(i).
\]
Thus an ordered $n$-tuple is simply a function from an index set.
\end{remark}

\begin{definition}[Operations in $\mathbb{F}^n$]
For $v,w \in \mathbb{F}^n$ and $\alpha \in \mathbb{F}$, define:
\[
(v+w)(i) := v(i) + w(i),
\]
\[
(\alpha v)(i) := \alpha \, v(i),
\]
for each $i \in \{1,\dots,n\}$.
\end{definition}

\begin{remark}[Pointwise structure]
Vector addition and scalar multiplication in $\mathbb{F}^n$
are defined \emph{pointwise}.
All vector space axioms follow from the corresponding field axioms
applied coordinatewise.
\end{remark}

\begin{remark}[Structural Insight]
When Axler writes
\[
v = (v_1,\dots,v_n),
\]
he is implicitly identifying $v$ with a function
\[
v : \{1,\dots,n\} \to \mathbb{F}.
\]
Thus
\[
\mathbb{F}^n = \{ f : \{1,\dots,n\} \to \mathbb{F} \}.
\]
This function viewpoint makes the structure of $\mathbb{F}^n$
completely rigorous and conceptually clean.
\end{remark}

% ---------------------------------------------------------
\subsubsection{The Function-Space Viewpoint}

\begin{remark}
Our next example of a vector space involves a set of functions.
Just as $\mathbb{F}^n$ consists of functions from a finite index set into $\mathbb{F}$,
we may replace that finite index set with any set $S$ whatsoever.
The resulting space $\mathbb{F}^S$ is again a vector space,
and this construction subsumes $\mathbb{F}^n$ as a special case.
\end{remark}

\begin{definition}[Set Exponentiation]
Let $X$ and $Y$ be sets.
Define
\[
Y^X := \{ f : X \to Y \},
\]
the set of all functions from $X$ to $Y$.
\end{definition}

\begin{remark}
If $|X| = n$ and $|Y| = m$ are finite, then
\[
|Y^X| = m^n.
\]
Thus the notation $Y^X$ agrees with the rules of cardinal arithmetic.
\end{remark}

\begin{definition}[$\mathbb{F}^S$ — the function space over a set]
Let $\mathbb{F}$ be a field and let $S$ be a set.
Define
\[
\mathbb{F}^S := \{ f : S \to \mathbb{F} \},
\]
the set of all functions from $S$ to $\mathbb{F}$.

For $f, g \in \mathbb{F}^S$, define \textbf{addition} by
\[
(f + g)(x) := f(x) + g(x)
\quad \text{for all } x \in S.
\]

For $\lambda \in \mathbb{F}$ and $f \in \mathbb{F}^S$, define \textbf{scalar multiplication} by
\[
(\lambda f)(x) := \lambda f(x)
\quad \text{for all } x \in S.
\]
\end{definition}

\begin{example}[$\mathbb{F}^S$ is a vector space]
Let $\mathbb{F}$ be a field and let $S$ be a set.
With the addition and scalar multiplication defined above,
$\mathbb{F}^S$ is a vector space over $\mathbb{F}$.

The zero vector is the function $\mathbf{0} : S \to \mathbb{F}$ defined by
$\mathbf{0}(x) = 0$ for all $x \in S$,
and the additive inverse of $f$ is the function $-f$ defined by
$(-f)(x) = -f(x)$ for all $x \in S$.
All vector space axioms follow from the corresponding field axioms applied pointwise.
\end{example}

\begin{definition}[Coordinate space as a function space]
Let $\mathbb{F}$ be a field and let
\[
[n] := \{1,\dots,n\}.
\]
Then
\[
\mathbb{F}^n \;=\; \mathbb{F}^{[n]}
\;=\;
\{ f : [n] \to \mathbb{F} \}.
\]
\end{definition}

\begin{remark}
An element $v \in \mathbb{F}^n$ is therefore a function
\[
v : [n] \to \mathbb{F}.
\]
The ordered tuple notation
\[
v = (v_1,\dots,v_n)
\]
means precisely that
\[
v(i) = v_i.
\]
Thus an $n$-tuple is simply a function from a finite index set.
\end{remark}

\begin{definition}[Pointwise operations]
Let $v,w \in \mathbb{F}^n$ and $\alpha \in \mathbb{F}$.
Define
\[
(v+w)(i) := v(i) + w(i),
\qquad
(\alpha v)(i) := \alpha\, v(i),
\]
for each $i \in [n]$.
\end{definition}

\begin{remark}
Vector addition and scalar multiplication in $\mathbb{F}^n$
are defined \emph{pointwise}.
The vector space axioms follow from the field axioms
applied at each index $i$.
\end{remark}

\begin{remark}[Structural Insight]
The notation
\[
\mathbb{F}^n = \mathbb{F}^{[n]}
\]
reveals that finite-dimensional coordinate spaces
are function spaces on finite index sets.

More generally, for any set $S$,
\[
\mathbb{F}^S := \{ f : S \to \mathbb{F} \}
\]
is a vector space under pointwise operations.

Thus finite-dimensional linear algebra is the special case
of function spaces where the index set is finite.
\end{remark}

\begin{remark}[Conceptual Unification]
The passage
\[
\mathbb{F}^n
\quad\longrightarrow\quad
\mathbb{F}^{\mathbb{N}}
\quad\longrightarrow\quad
\mathbb{F}^S
\]
shows that:
\begin{itemize}
\item Finite tuples are functions on finite sets.
\item Sequences are functions on $\mathbb{N}$.
\item General function spaces are vector spaces indexed by arbitrary sets.
\end{itemize}

This viewpoint unifies finite-dimensional linear algebra
with infinite-dimensional vector spaces and
functional analysis.
\end{remark}

\begin{definition}[Zero vector]
Let $V$ be a vector space over a field $\mathbb{F}$.
The \emph{zero vector} of $V$, denoted $\mathbf{0}$, is the additive identity
of the abelian group $(V,+)$; that is, the unique element $\mathbf{0} \in V$
such that
\[
v + \mathbf{0} = v
\quad \text{for all } v \in V.
\]
\end{definition}

\begin{remark}[Zero vector in $\mathbb{F}^S$]
In the function space $\mathbb{F}^S$, the zero vector is the function
\[
\mathbf{0} : S \to \mathbb{F}
\quad \text{defined by} \quad
\mathbf{0}(x) = 0
\text{ for all } x \in S.
\]
\end{remark}

% ---------------------------------------------------------
\subsubsection{The Space $\mathbb{F}^{\infty}$}

\begin{definition}[$\mathbb{F}^{\infty}$ — the space of sequences]
Let $\mathbb{F}$ be a field.
Define
\[
\mathbb{F}^{\infty}
:=
\{ f : \mathbb{N} \to \mathbb{F} \}.
\]
Equivalently,
\[
\mathbb{F}^{\infty}
=
\{ (x_1,x_2,\dots) : x_k \in \mathbb{F} \text{ for all } k \in \mathbb{N} \}.
\]
An element of $\mathbb{F}^{\infty}$ is called a \emph{sequence} of elements of $\mathbb{F}$.
\end{definition}

\begin{definition}[Pointwise operations on $\mathbb{F}^{\infty}$]
Let $x=(x_1,x_2,\dots)$ and $y=(y_1,y_2,\dots)$ be elements of $\mathbb{F}^{\infty}$,
and let $\lambda \in \mathbb{F}$.

Define addition by
\[
x+y := (x_1+y_1,\; x_2+y_2,\; \dots),
\]
and define scalar multiplication by
\[
\lambda x := (\lambda x_1,\; \lambda x_2,\; \dots).
\]
Equivalently, in function notation:
\[
(x+y)(k) := x(k)+y(k),
\qquad
(\lambda x)(k) := \lambda x(k),
\quad k \in \mathbb{N}.
\]
\end{definition}

\begin{remark}
With the pointwise operations defined above,
$\mathbb{F}^{\infty}$ is a vector space over $\mathbb{F}$.
All vector space axioms follow from the corresponding field axioms,
applied coordinatewise.
\end{remark}

\begin{remark}[Structural Perspective]
The space $\mathbb{F}^{\infty}$ is simply the function space
\[
\mathbb{F}^{\mathbb{N}},
\]
the set of all $\mathbb{F}$-valued functions on $\mathbb{N}$.
Thus:
\[
\mathbb{F}^n = \mathbb{F}^{\{1,\dots,n\}}
\quad\text{and}\quad
\mathbb{F}^{\infty} = \mathbb{F}^{\mathbb{N}}
\]
are instances of the same construction:
vector spaces of functions indexed by a set.
\end{remark}

% ---------------------------------------------------------
\subsubsection{Example: $\mathbb{F}^S$ is a Vector Space}

\begin{example}[$\mathbb{F}^S$ is a vector space]
Let $\mathbb{F}$ be a field and let $S$ be a nonempty set.
With addition and scalar multiplication defined pointwise by
\[
(f+g)(x) := f(x) + g(x),
\qquad
(\lambda f)(x) := \lambda f(x),
\]
for all $f,g \in \mathbb{F}^S$, $\lambda \in \mathbb{F}$, and $x \in S$,
the set $\mathbb{F}^S$ is a vector space over $\mathbb{F}$.
\end{example}

\begin{remark}[Additive identity]
The additive identity of $\mathbb{F}^S$ is the function
\[
\mathbf{0} : S \to \mathbb{F}
\quad \text{defined by} \quad
\mathbf{0}(x) = 0
\text{ for all } x \in S.
\]
\end{remark}

\begin{remark}[Additive inverse]
For $f \in \mathbb{F}^S$, the additive inverse of $f$ is the function
\[
- f : S \to \mathbb{F}
\quad \text{defined by} \quad
(-f)(x) = -\,f(x)
\text{ for all } x \in S.
\]
\end{remark}

\begin{remark}[Why the axioms hold]
All vector space axioms follow from the field axioms,
applied pointwise. For example,
\[
((f+g)+h)(x)
=
f(x)+g(x)+h(x)
=
(f+(g+h))(x),
\]
so addition is associative because addition in $\mathbb{F}$ is associative.
Every other axiom follows similarly.
\end{remark}























