% =========================================================
% Proof Techniques — Notes Assembly
% =========================================================
% =========================================================
% Tao Analysis I — Chapter 2 Reference Sheet
% Available axioms, definitions, and lemmas
% =========================================================

\subsection*{Peano Axioms}

\begin{itemize}
  \item[\textbf{P1.}] $0 \in \mathbb{N}$
  \item[\textbf{P2.}] $n \in \mathbb{N} \Rightarrow n\pp \in \mathbb{N}$
  \item[\textbf{P3.}] $n\pp \neq 0$ for all $n \in \mathbb{N}$
  \item[\textbf{P4.}] $n\pp = m\pp \Rightarrow n = m$ \quad (successor is injective)
  \item[\textbf{P5.}] \emph{Induction axiom}: if $P(0)$ holds and
    $P(n) \Rightarrow P(n\pp)$ for all $n \in \mathbb{N}$,
    then $P(n)$ holds for all $n \in \mathbb{N}$.
\end{itemize}

\subsection*{Definition of Addition (Recursive)}

\begin{itemize}
  \item[\textbf{A1.}] $0 + m := m$
  \item[\textbf{A2.}] $(n\pp) + m := (n + m)\pp$
\end{itemize}

\noindent\textit{Note}: recursion is on the \textbf{left} argument only.
$n + m$ is defined by recursing on $n$, not $m$.

\subsection*{Available Results}

\begin{itemize}
  \item[\textbf{L2.2.2.}] $n + 0 = n$ \quad for all $n \in \mathbb{N}$
  \item[\textbf{L2.2.3.}] $n + (m\pp) = (n + m)\pp$ \quad
    for all $n, m \in \mathbb{N}$
  \item[\textbf{P2.2.4.}] $n + m = m + n$ \quad
    for all $n, m \in \mathbb{N}$ \quad (commutativity)
\end{itemize}





\begin{tcolorbox}[
  colback=gray!6, colframe=gray!40, arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic $\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ \textbf{Proof Techniques}
$\;\to\;$ Real Analysis $\;\to\;$ Algebraic Structures
$\;\to\;$ Linear Algebra $\;\to\;$ $\cdots$
\end{center}
\medskip
\noindent\textbf{How we got here.}
Logic gave us the language of valid inference and sets gave us the
objects we reason about. This chapter asks the operational question:
how do we build a valid mathematical argument from start to finish?

\medskip
\noindent\textbf{What this chapter builds.}
A complete architecture for proof writing: how to read a statement
and identify its proof strategy immediately; how to execute that
strategy step by step; how to use induction to reason about
well-ordered and recursive structures; and how to deploy specific
algebraic tactics when progress stalls.

\medskip
\noindent\textbf{Where this leads.}
Every proof in every subsequent chapter uses this toolkit.
Epsilon-delta analysis uses direct proof and contradiction.
Algebraic structure proofs use the uniqueness and
satisfy-and-cite patterns. Induction appears throughout
number theory, algebra, and combinatorics.
\end{tcolorbox}
\vspace{1em}

\subsection*{Structural Roadmap}

This chapter is organized in six sections of increasing specificity.

\begin{center}
\textbf{Architecture $\longrightarrow$ Construction
$\longrightarrow$ Structures $\longrightarrow$ Induction
$\longrightarrow$ Tactics $\longrightarrow$ Reference}
\end{center}

Sections 1--4 should be read sequentially on first pass.
Sections 5--6 are reference material to consult when working a
proof and progress stalls.

\begin{remark}[Four distinct questions]
\textbf{Architecture} answers: what kind of proof is this and what
skeleton should it have?
\textbf{Construction} answers: how do I execute that skeleton
line by line?
\textbf{Structures} answers: what do the standard proof patterns
look like in practice?
\textbf{Tactics} answers: once a proof is in progress,
what specific moves are available?
These are four different questions. Conflating them is the most
common source of proof-writing paralysis.
\end{remark}

\input{volume-i/proof-techniques/notes/architecture/index}
\input{volume-i/proof-techniques/notes/construction/index}
\input{volume-i/proof-techniques/notes/structures/index}
\input{volume-i/proof-techniques/notes/induction/index}
\input{volume-i/proof-techniques/notes/tactics/index}
\input{volume-i/proof-techniques/notes/reference/index}
