% Fragment (no \documentclass)

\subsubsection*{MAG-P1-C01-WS-001 (10 problems)}\n\hypertarget{MAG-P1-C01-WS-001}{}\n\n\noindent\textbf{Navigation: }
\hyperlink{P1-MAG-C01}{Back to chapter}\n\n\begin{enumerate}
\item \textbf{Inclusion is partial order.} Prove that $\subseteq$ is a partial order on $\mathcal{P}(X)$ for any set $X$. (\hyperlink{MAG-P1-C01-001}{proof})
\item \textbf{Equality is equivalence relation.} Prove that equality is an equivalence relation. (\hyperlink{MAG-P1-C01-002}{proof})
\item \textbf{Power set poset.} Prove that $(\mathcal{P}(X),\subseteq)$ is a partially ordered set. (\hyperlink{MAG-P1-C01-003}{proof})
\item \textbf{Preimages preserve unions.} Let $f:X\to Y$. Prove that preimages preserve unions. (\hyperlink{MAG-P1-C01-004}{proof})
\item \textbf{Preimages preserve intersections.} Let $f:X\to Y$. Prove that preimages preserve intersections. (\hyperlink{MAG-P1-C01-005}{proof})
\item \textbf{Preimages preserve complements.} Let $f:X\to Y$. Prove that preimages preserve complements. (\hyperlink{MAG-P1-C01-006}{proof})
\item \textbf{Intersection of equivalence relations?.} Give an example showing the intersection of two equivalence relations need not be an equivalence relation, or prove it always is (whichever is true). (\hyperlink{MAG-P1-C01-007}{proof})
\item \textbf{Composition not symmetric.} Give an example showing that the composition of relations need not be symmetric. (\hyperlink{MAG-P1-C01-008}{proof})
\item \textbf{Uniqueness of limits.} Prove uniqueness of limits in a metric space. (\hyperlink{MAG-P1-C01-009}{proof})
\item \textbf{Convergent implies Cauchy.} Prove that every convergent sequence in a metric space is Cauchy. (\hyperlink{MAG-P1-C01-010}{proof})
\end{enumerate}

\subsubsection*{Proofs (included)}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_001.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_002.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_003.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_004.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_005.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_006.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_007.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_008.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_009.tex}
\input{proofs/phase1/magnus/ch01/MAG_P1_C01_010.tex}
