% =========================================================
% Set Algebras — Structural Roadmap
% File: set-algebras/set-algebras.tex
% =========================================================

\subsection{Set Algebras}

% =========================================================
% Structural Roadmap
% =========================================================
\subsubsection*{Structural Roadmap}

The study of set algebras develops a structural theory
of families of subsets of a fixed space.
Rather than focusing on individual elements,
we investigate the \emph{stability properties}
of collections of sets under recombination.

The central question is:

\begin{center}
\emph{Under which operations is a family of sets closed?}
\end{center}

Different closure assumptions give rise to
distinct algebraic structures on $2^X$.

Each major topic follows the architecture:
\begin{center}
\textbf{Definitions $\longrightarrow$ Structural Consequences $\longrightarrow$ Hierarchy of Systems}
\end{center}

The global progression is:

\begin{enumerate}
  \item Families of sets and closure operations
  \item Finite closure structures:
        rings, algebras (fields), Boolean algebras
  \item Countable closure structures:
        $\sigma$-rings and $\sigma$-algebras
  \item Generated systems and minimal closure
  \item Auxiliary systems (e.g.\ $\pi$-systems, Dynkin systems)
  \item Structural comparison of closure regimes
\end{enumerate}

\vspace{1em}

\begin{remark}
Set algebras form a theory of
\emph{closure under operations}.
They classify families of sets according to how
stable they are under union, intersection,
complement, difference, and limit-type constructions.
\end{remark}

\begin{remark}
Finite closure produces algebraic stability;
countable closure prepares compatibility with
limit processes.
These structures will later appear naturally in
topology, measure theory, probability,
and Boolean algebra,
but they are studied here independently
as algebraic systems of subsets.
\end{remark}

% =========================================================
% I. Systems of Sets
% =========================================================

\input{intro-set-algebras/notes/notes-set-algebras}
\input{intro-set-algebras/notes/notes-power-set-and-characteristic-functions}
% =========================================================
% Future Sections (Expandable Architecture)
% =========================================================

% Uncomment as content expands:

% \input{intro-set-algebras/notes-finite-structures}
% \input{intro-set-algebras/notes-boolean-algebras}
% \input{intro-set-algebras/notes-sigma-structures}
% \input{intro-set-algebras/notes-generated-systems}
% \input{intro-set-algebras/notes-auxiliary-systems}
% \input{intro-set-algebras/notes-applications}

% =========================================================
% Logical Positioning
% =========================================================

\begin{remark}[Structural Position]
This section isolates the algebraic behavior
of systems of sets without committing
to any particular application.

Topology will later arise from arbitrary unions
and finite intersections;
measure theory will arise from
$\sigma$-algebras equipped with additive functionals.
Both are specializations of the general closure
framework developed here.
\end{remark}