% =========================================================
% Natural Numbers — Chapter
% =========================================================
\chapter{Natural Numbers ($\mathbb{N}$)}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ \textbf{Natural Numbers ($\mathbb{N}$)}
$\;\to\;$ Integers ($\mathbb{Z}$)
$\;\to\;$ Rationals ($\mathbb{Q}$)
$\;\to\;$ Reals ($\mathbb{R}$)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The Peano axioms in Volume~I (Axiom Systems) constructed $\mathbb{N}$
from scratch and proved all arithmetic properties by induction.
This chapter collects the resulting properties of $\mathbb{N}$ ---
arithmetic, order, well-ordering, and divisibility --- as a
self-contained reference.

\medskip
\noindent\textbf{What this chapter will build.}
Properties of addition and multiplication, the well-ordering principle,
strong induction, divisibility, primes, and the fundamental theorem
of arithmetic.

\medskip
\noindent\textbf{Where this leads.}
The integers extend $\mathbb{N}$ by adjoining additive inverses.
Every fact about $\mathbb{N}$ carries forward into $\mathbb{Z}$,
$\mathbb{Q}$, and $\mathbb{R}$.
\end{tcolorbox}

\vspace{1em}

\begin{tcolorbox}[
  colback=gray!6, colframe=gray!40, arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Navigation Note}},
  fonttitle=\small\bfseries
]
\noindent
The \textbf{axiomatic construction of $\mathbb{N}$} via the Peano Axioms lives in
Volume~I under \textit{Axiom Systems}.
This chapter collects the resulting \emph{properties, arithmetic, and order
theory of $\mathbb{N}$} as a self-contained reference.
\end{tcolorbox}

\section{Notes}
\textit{To be populated. Planned topics: successor function, addition,
multiplication, order, well-ordering principle, strong induction.}

\section{Proofs}
\textit{To be populated.}

\section{Capstone}
\textit{To be populated.}
