% =========================================================
% Rational Numbers
% =========================================================
\chapter{Rational Numbers ($\mathbb{Q}$)}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
$\mathbb{N}$
$\;\to\;$ Integers ($\mathbb{Z}$)
$\;\to\;$ \textbf{Rationals ($\mathbb{Q}$)}
$\;\to\;$ Real Numbers ($\mathbb{R}$)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The integers gave us additive inverses, so subtraction became internal to the number system.
Division still fails: for example, $1 \div 2$ is not an integer.  The rational numbers adjoin
multiplicative inverses for all nonzero integers.

\medskip
\noindent\textbf{What this chapter develops.}
We construct $\mathbb{Q}$ from $\mathbb{Z}$ using equivalence classes of ordered pairs, prove that
$\mathbb{Q}$ is an ordered field, study its density and approximation properties, and then see why
$\mathbb{Q}$ is still incomplete.

\medskip
\noindent\textbf{Where this leads.}
The rational numbers are already rich enough to support most of arithmetic and much of algebra, but
not rich enough to capture all limits.  That failure motivates the construction of $\mathbb{R}$.
\end{tcolorbox}

\section{Notes}

\input{volume-ii/rationals/notes/figure-q-roadmap}
\input{volume-ii/rationals/notes/notes-rational}
\input{volume-ii/rationals/notes/figure-constructing-rationals}
\input{volume-ii/rationals/notes/notes-field-order-axioms}
\input{volume-ii/rationals/notes/notes-density-of-rationals}
\input{volume-ii/rationals/notes/notes-mediant}
\input{volume-ii/rationals/notes/figure-stern-brocot}
\input{volume-ii/rationals/notes/figure-mediant-archimedean}
\input{volume-ii/rationals/notes/notes-rational-approximation}
\input{volume-ii/rationals/notes/figure-rational-approximation}
\input{volume-ii/rationals/notes/figure-dirichlet-aproximation}
\input{volume-ii/rationals/notes/rational-cauchy}
\input{volume-ii/rationals/notes/notes-gaps-in-rationals}

\section{Proofs}
\input{volume-ii/rationals/proofs/index.tex.tex}

\section{Capstone}
\textit{To be populated.}
