% =========================================================
% Proof Grep Stubs — Rational Numbers (Chapter 10)
% Source: Learning_Real_Analysis (76).pdf
% Purpose: Combined proof-stub file for theorem-like statements
% Mode: standard stubs only (no completed proofs)
% =========================================================

\section*{Proof Grep — Rational Numbers}

This file collects proof stubs for the theorem-like statements in Chapter~10
of the current Rational Numbers PDF.  The statements have been arranged in a
\emph{dependency-aware order} so that earlier stubs do not rely on later ones.

\begin{remark*}[Dependency order used]
The proof stubs below are ordered so that structural results about the quotient
construction and ordered-field consequences appear before density, mediants,
Farey phenomena, rational approximation, Cauchy-sequence results, and finally
limiting-process results involving Stern--Brocot refinement.
\end{remark*}

\begin{remark*}[Duplicate formulations consolidated]
The current PDF contains three duplicate theorem/corollary pairs:
\begin{itemize}
  \item Theorem~10.25 and Theorem~10.40 both state $1 > 0$.
  \item Corollary~10.26 and Corollary~10.41 both state $n > 0$ for all $n \in \mathbb{N}$.
  \item Theorem~10.33 and Theorem~10.49 are both mediant inequalities.
\end{itemize}
To avoid redundant stubs, this file keeps the canonical occurrences
Theorem~10.25, Corollary~10.26, and Theorem~10.49.
\end{remark*}

\subsection*{Theorem 10.6 — The relation $\sim$ is an equivalence relation on $W$}
\label{prf:grep-rational-equivalence-relation}

\begin{proof}
\Claim The relation $\sim$ on $W$ defined by
\[
(a,b) \sim (c,d) \iff ad = bc
\]
is an equivalence relation on $W$.

\textbf{Given.} The set $W = \mathbb{Z} \times (\mathbb{Z}\setminus\{0\})$ and the relation above.

\textbf{Goal.} Show that $\sim$ is reflexive, symmetric, and transitive.

\textbf{Strategy.} Verify the three defining properties of an equivalence relation directly from the defining equation $ad=bc$.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a direct verification proof.  The reflexive and symmetric parts are
immediate algebraic identities; the transitive part is the only nontrivial step
and should be handled by chaining two cross-multiplication equalities.
\end{remark*}

\begin{remark*}[Dependencies]
Uses only the definition of $W$, the definition of $\sim$, and basic arithmetic
of the integers.
\end{remark*}


\subsection*{Theorem 10.21 — For every $a \in F$, $a^2 \ge 0$}
\label{prf:grep-squares-nonnegative}

\begin{proof}
\Claim For every $a \in F$,
\[
a^2 \ge 0.
\]

\textbf{Given.} An ordered field $F$ presented via a positive cone $P$.

\textbf{Goal.} Show that every square is positive or zero.

\textbf{Strategy.} Apply trichotomy to $a$ and split into the cases $a \in P$, $a=0$, and $-a \in P$.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a three-case proof obtained by unpacking the positive-cone axioms.
The key observation is that if $-a$ is positive, then $(-a)(-a)=a^2$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Definition~10.19 (positive cone), especially trichotomy and closure under
multiplication.
\end{remark*}


\subsection*{Theorem 10.25 — In an ordered field, $1 > 0$}
\label{prf:grep-one-positive}

\begin{proof}
\Claim In an ordered field,
\[
1 > 0.
\]

\textbf{Given.} An ordered field (or equivalently a field with a positive cone).

\textbf{Goal.} Show that the multiplicative identity is positive.

\textbf{Strategy.} Use trichotomy on $1$.  Eliminate the possibilities $1=0$ and $-1>0$.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a contradiction-by-cases proof.  The case $1=0$ contradicts the field
axioms, and the case $-1>0$ collapses under multiplication.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Definition~10.19 (positive cone), the field axiom $1 \neq 0$, and closure
of positive elements under multiplication.
\end{remark*}


\subsection*{Theorem 10.22 — If $a>0$, then $a^{-1}>0$}
\label{prf:grep-positive-inverse}

\begin{proof}
\Claim If $a>0$, then
\[
a^{-1} > 0.
\]

\textbf{Given.} An ordered field and an element $a$ with $a>0$.

\textbf{Goal.} Show that the multiplicative inverse of $a$ is positive.

\textbf{Strategy.} Suppose the inverse were not positive and derive a contradiction by multiplying by a positive element.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a contradiction proof.  The expected contradiction is that a positive
number multiplied by an allegedly negative inverse would force $-1$ to be
positive, contradicting Theorem~10.25.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.25, Definition~10.19, and the field axiom that every nonzero
element has a multiplicative inverse.
\end{remark*}


\subsection*{Theorem 10.23 — If $a<b$ and $c>0$, then $ac<bc$}
\label{prf:grep-multiply-by-positive}

\begin{proof}
\Claim If $a<b$ and $c>0$, then
\[
ac < bc.
\]

\textbf{Given.} Elements $a,b,c$ in an ordered field with $a<b$ and $c>0$.

\textbf{Goal.} Show that multiplication by a positive element preserves strict inequality.

\textbf{Strategy.} Translate $a<b$ into positivity of $b-a$, then multiply by $c$ and translate back.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a direct proof entirely inside the positive-cone language:
$b-a \in P$, $c \in P$, hence $c(b-a) \in P$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Definition~10.19 and closure of the positive cone under multiplication.
\end{remark*}


\subsection*{Theorem 10.24 — If $a<b$ and $c<0$, then $ac>bc$}
\label{prf:grep-multiply-by-negative}

\begin{proof}
\Claim If $a<b$ and $c<0$, then
\[
ac > bc.
\]

\textbf{Given.} Elements $a,b,c$ in an ordered field with $a<b$ and $c<0$.

\textbf{Goal.} Show that multiplication by a negative element reverses strict inequality.

\textbf{Strategy.} Replace $c$ by $-c>0$, apply Theorem~10.23, and then rewrite the resulting inequality.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a reduction proof: first convert the negative multiplier into a
positive one, then appeal to the previous theorem.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.23 and Definition~10.19.
\end{remark*}


\subsection*{Corollary 10.26 — For every natural number $n$, $n>0$}
\label{prf:grep-natural-numbers-positive}

\begin{proof}
\Claim For every natural number $n$,
\[
n > 0.
\]

\textbf{Given.} An ordered field and a natural number $n$ regarded as a finite sum of copies of $1$.

\textbf{Goal.} Show that every natural number is positive.

\textbf{Strategy.} Combine Theorem~10.25 with closure of the positive cone under addition.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a short consequence proof.  Once $1>0$ is known, every finite sum of
copies of $1$ is positive.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.25 and Definition~10.19.
\end{remark*}


\subsection*{Theorem 10.42 — Archimedean property of $\mathbb{Q}$}
\label{prf:grep-archimedean-q}

\begin{proof}
\Claim For every $x \in \mathbb{Q}$ there exists $n \in \mathbb{N}$ such that
\[
n > x.
\]
Equivalently, for every $x>0$ there exists $n \in \mathbb{N}$ such that
\[
\frac{1}{n} < x.
\]

\textbf{Given.} A rational number $x$.

\textbf{Goal.} Produce a natural number that exceeds $x$.

\textbf{Strategy.} Write $x$ in fraction form and choose a natural number large enough relative to its numerator and denominator.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a construction proof specific to $\mathbb{Q}$.  The natural approach is
to write $x=a/b$ and choose an explicit natural number in terms of $a$ and $b$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses the construction of $\mathbb{Q}$, Corollary~10.26, and basic arithmetic of
integers and rational numbers.
\end{remark*}


\subsection*{Theorem 10.44 — Density of $\mathbb{Q}$ in $\mathbb{R}$}
\label{prf:grep-density-q-in-r}

\begin{proof}
\Claim If $x,y \in \mathbb{R}$ and $x<y$, then there exists $q \in \mathbb{Q}$ such that
\[
x < q < y.
\]

\textbf{Given.} Real numbers $x<y$.

\textbf{Goal.} Find a rational number strictly between them.

\textbf{Strategy.} Use the Archimedean property to choose $n$ with $1/n < y-x$, then locate an integer $m$ with $x < m/n \le x+1/n$.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is the standard Archimedean-density argument.  The key move is to pass
from interval length to a denominator and then select the least integer above
$nx$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.42 and the order structure of the real numbers.
\end{remark*}


\subsection*{Corollary 10.45 — Between any two distinct real numbers there are infinitely many rationals}
\label{prf:grep-infinitely-many-rationals}

\begin{proof}
\Claim Between any two distinct real numbers there are infinitely many rational numbers.

\textbf{Given.} Real numbers $x<y$.

\textbf{Goal.} Show that the interval $(x,y)$ contains infinitely many rationals.

\textbf{Strategy.} Apply Theorem~10.44 repeatedly to smaller subintervals.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is an iteration proof.  Once one rational is found, the density theorem can
be re-applied to the left and right subintervals or to a recursively chosen chain
of subintervals.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.44.
\end{remark*}


\subsection*{Theorem 10.49 — Mediant inequality}
\label{prf:grep-mediant-inequality}

\begin{proof}
\Claim If
\[
\frac{a}{b} < \frac{c}{d}
\quad\text{with } b,d>0,
\]
then
\[
\frac{a}{b} < \frac{a+c}{b+d} < \frac{c}{d}.
\]

\textbf{Given.} Two rational numbers with positive denominators and strict order.

\textbf{Goal.} Show that the mediant lies strictly between them.

\textbf{Strategy.} Start from the cross-multiplied inequality $ad<bc$ and manipulate it separately to obtain the left-hand and right-hand bounds.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a two-inequality algebra proof.  Each side comes from adding a suitable
common term to $ad<bc$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses only the order properties of rational numbers and basic algebraic
manipulation.
\end{remark*}


\subsection*{Corollary 10.50 — Density of $\mathbb{Q}$ within itself}
\label{prf:grep-density-q-within-q}

\begin{proof}
\Claim If $r,s \in \mathbb{Q}$ with $r<s$, then there exists $t \in \mathbb{Q}$ such that
\[
r < t < s.
\]

\textbf{Given.} Rational numbers $r<s$.

\textbf{Goal.} Produce a rational number strictly between them.

\textbf{Strategy.} Write $r$ and $s$ with positive denominators and apply the mediant inequality.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is the immediate corollary of Theorem~10.49.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.49 and the fact that rational numbers can be represented with
positive denominators.
\end{remark*}


\subsection*{Theorem 10.61 — Farey Neighbor Theorem}
\label{prf:grep-farey-neighbor-theorem}

\begin{proof}
\Claim Let
\[
\frac{a}{b} < \frac{c}{d}
\]
be two consecutive fractions in the Farey sequence $F_n$. Then
\[
bc-ad = 1.
\]

\textbf{Given.} Consecutive fractions in a Farey sequence.

\textbf{Goal.} Show that their determinant difference is exactly $1$.

\textbf{Strategy.} Assume the determinant difference is at least $2$, then construct a reduced fraction with denominator at most $n$ lying strictly between them, contradicting consecutiveness.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a contradiction proof.  The heart of the argument is that if the gap is
``too large'' arithmetically, then an intermediate Farey fraction must exist.
\end{remark*}

\begin{remark*}[Dependencies]
Uses the definition of the Farey sequence and elementary properties of reduced
fractions and inequalities.
\end{remark*}


\subsection*{Theorem 10.65 — Gap Between Farey Neighbors}
\label{prf:grep-farey-gap}

\begin{proof}
\Claim Let
\[
\frac{a}{b} < \frac{c}{d}
\]
be consecutive fractions in a Farey sequence. Then
\[
\frac{c}{d} - \frac{a}{b} = \frac{1}{bd}.
\]

\textbf{Given.} Consecutive Farey neighbors.

\textbf{Goal.} Compute their exact spacing.

\textbf{Strategy.} Subtract the two fractions and then use the determinant identity from Theorem~10.61.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a one-line computation once the Farey determinant condition is known.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.61 and direct fraction subtraction.
\end{remark*}


\subsection*{Theorem 10.52 — Farey neighbor determinant condition}
\label{prf:grep-farey-neighbor-determinant-condition}

\begin{proof}
\Claim Suppose
\[
\frac{a}{b} < \frac{c}{d}
\quad\text{with } \gcd(a,b)=\gcd(c,d)=1,\; b,d>0.
\]
If
\[
bc-ad = 1,
\]
then the two fractions are neighbors in a Farey sequence, and their mediant
\[
\frac{a+c}{b+d}
\]
is already in reduced form.

\textbf{Given.} Two reduced fractions with positive denominators satisfying $bc-ad=1$.

\textbf{Goal.} Prove both the Farey-neighbor conclusion and the reduced-form conclusion for the mediant.

\textbf{Strategy.} For the neighbor claim, argue that any hypothetical reduced fraction strictly between them would force a contradiction with the determinant condition.  For reducedness of the mediant, show that a common divisor would have to divide $bc-ad=1$.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a two-part proof: a best-approximation/neighbor argument and an
arithmetic gcd argument.
\end{remark*}

\begin{remark*}[Dependencies]
Uses the definition of Farey sequence, basic gcd properties, and the arithmetic
identity $bc-ad=1$.
\end{remark*}


\subsection*{Theorem 10.62 — Mediant Property of Farey Neighbors}
\label{prf:grep-mediant-farey-neighbors}

\begin{proof}
\Claim If
\[
\frac{a}{b},\frac{c}{d}
\]
are consecutive fractions in a Farey sequence, then the mediant
\[
\frac{a+c}{b+d}
\]
lies strictly between them.

\textbf{Given.} Consecutive Farey neighbors.

\textbf{Goal.} Show that their mediant sits strictly between them.

\textbf{Strategy.} Apply the general mediant inequality to these two fractions.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a direct corollary of the general mediant inequality.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.49.
\end{remark*}


\subsection*{Lemma 10.70 — Mediant Gap Lemma}
\label{prf:grep-mediant-gap-lemma}

\begin{proof}
\Claim If
\[
\frac{a}{b} < \frac{c}{d},
\]
then the mediant
\[
\frac{a+c}{b+d}
\]
lies between them, and the gaps satisfy
\[
\frac{a+c}{b+d} - \frac{a}{b} = \frac{1}{b(b+d)}
\]
and
\[
\frac{c}{d} - \frac{a+c}{b+d} = \frac{1}{d(b+d)}.
\]

\textbf{Given.} Two ordered rational numbers with positive denominators.

\textbf{Goal.} Compute the two subgap lengths created by inserting the mediant.

\textbf{Strategy.} First establish that the mediant lies between the endpoints.  Then subtract fractions directly and simplify.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is an explicit computation proof.  The interval-refinement content comes
from exact arithmetic rather than from an abstract limit argument.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.49 and routine fraction subtraction.
\end{remark*}


\subsection*{Theorem 10.54 — Stern--Brocot enumeration}
\label{prf:grep-stern-brocot-enumeration}

\begin{proof}
\Claim Every positive rational number appears exactly once in reduced form in the Stern--Brocot tree.

\textbf{Given.} The Stern--Brocot tree generated from $0/1$ and $1/0$ by repeated mediants.

\textbf{Goal.} Show both existence and uniqueness of appearance for every positive rational.

\textbf{Strategy.} Separate the proof into two parts: (1) every positive rational can be reached by a finite mediant-search path; (2) no reduced rational appears twice.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a two-part structural proof.  The existence part is usually handled by a
finite search or Euclidean-algorithm argument; the uniqueness part relies on the
determinant/neighbor structure of Stern--Brocot intervals.
\end{remark*}

\begin{remark*}[Dependencies]
Uses the mediant inequality, Farey/Stern--Brocot determinant phenomena, and
basic properties of reduced fractions.
\end{remark*}


\subsection*{Theorem 10.72 — Rational approximation}
\label{prf:grep-rational-approximation}

\begin{proof}
\Claim For every real number $x$ and every $\varepsilon>0$, there exists $r \in \mathbb{Q}$ such that
\[
|x-r| < \varepsilon.
\]

\textbf{Given.} A real number $x$ and a positive tolerance $\varepsilon$.

\textbf{Goal.} Find a rational number within $\varepsilon$ of $x$.

\textbf{Strategy.} Apply density of $\mathbb{Q}$ to the interval $(x-\varepsilon, x+\varepsilon)$.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a one-step application of density.  The metric statement is obtained from
the order statement by choosing a symmetric interval around $x$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.44.
\end{remark*}


\subsection*{Theorem 10.76 — Dirichlet's Approximation Theorem}
\label{prf:grep-dirichlet-approximation}

\begin{proof}
\Claim For every real number $x$ there exist infinitely many rational numbers $p/q$ with $p \in \mathbb{Z}$ and $q \in \mathbb{N}$ such that
\[
\left|x-\frac{p}{q}\right| < \frac{1}{q^2}.
\]

\textbf{Given.} A real number $x$.

\textbf{Goal.} Produce infinitely many rational approximations satisfying the quadratic error bound.

\textbf{Strategy.} Use the pigeonhole principle on the fractional parts of $x,2x,\dots,(N+1)x$, derive an approximation of size $1/(qN)$, and then turn this into the desired $1/q^2$ bound for infinitely many choices.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is a combinatorial approximation proof driven by the pigeonhole principle.
The infinite-many conclusion usually comes from applying the finite-$N$ argument
for arbitrarily large $N$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses the pigeonhole principle, properties of fractional parts, and elementary
integer arithmetic.  It does not depend on later Cauchy- or completeness-based
results.
\end{remark*}


\subsection*{Theorem 10.91 — There is no rational number whose square is $2$}
\label{prf:grep-sqrt-two-irrational}

\begin{proof}
\Claim There is no rational number whose square is $2$.

\textbf{Given.} The assumption, for contradiction, that $\sqrt{2}=p/q$ for integers $p,q$ in lowest terms.

\textbf{Goal.} Derive a contradiction.

\textbf{Strategy.} Square the equation, prove that $p$ is even, then prove that $q$ is even, contradicting lowest terms.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is the classical parity contradiction proof of the irrationality of
$\sqrt{2}$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses reduced-form representatives of rational numbers and elementary parity
facts about integers.
\end{remark*}


\subsection*{Theorem 10.85 — Every convergent sequence in $\mathbb{Q}$ is a Cauchy sequence}
\label{prf:grep-convergent-implies-cauchy}

\begin{proof}
\Claim Every convergent sequence in $\mathbb{Q}$ is a Cauchy sequence.

\textbf{Given.} A sequence $(x_n)$ in $\mathbb{Q}$ converging to some $x \in \mathbb{Q}$.

\textbf{Goal.} Show that for every $\varepsilon>0$ there exists $N$ such that $m,n\ge N$ implies $|x_m-x_n|<\varepsilon$.

\textbf{Strategy.} Use the definition of convergence with tolerance $\varepsilon/2$ and apply the triangle inequality.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is the standard $\varepsilon/2$ triangle-inequality argument.
\end{remark*}

\begin{remark*}[Dependencies]
Uses the definition of convergence in $\mathbb{Q}$, the definition of Cauchy
sequence, and the triangle inequality.
\end{remark*}


\subsection*{Theorem 10.88 — There exist Cauchy sequences in $\mathbb{Q}$ that do not converge to a rational number}
\label{prf:grep-cauchy-not-convergent-in-q}

\begin{proof}
\Claim There exist Cauchy sequences in $\mathbb{Q}$ that do not converge to a rational number.

\textbf{Given.} A candidate sequence of rational approximations to $\sqrt{2}$, for example decimal truncations or another explicit approximating sequence.

\textbf{Goal.} Show that the sequence is Cauchy but has no rational limit.

\textbf{Strategy.} Split the proof into two parts: (1) verify the sequence is Cauchy; (2) prove that if it converged in $\mathbb{Q}$, its limit would have square $2$, contradicting Theorem~10.91.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is the first genuine incompleteness proof in the chapter.  The Cauchy part
is analytic; the nonconvergence part is arithmetic and should terminate in the
irrationality of $\sqrt{2}$.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.91 and the definition of Cauchy sequence.  Depending on the
chosen example, may also use a monotone-bound argument or direct decimal-error
estimates.
\end{remark*}


\subsection*{Theorem 10.96 — Stern--Brocot interval refinement}
\label{prf:grep-stern-brocot-interval-refinement}

\begin{proof}
\Claim Let $x>0$ be a real number.  Suppose one begins with two rational numbers
\[
\frac{a}{b} < x < \frac{c}{d}, \qquad b,d>0,
\]
and repeatedly inserts the mediant
\[
\frac{a+c}{b+d}.
\]
At each step, replace one endpoint by the mediant so that the new interval still contains $x$.  Then this process generates a nested sequence of rational intervals containing $x$.

\textbf{Given.} A target real number $x>0$ and an initial rational interval containing it.

\textbf{Goal.} Show that the iterative mediant-selection process yields nested rational intervals, each still containing $x$.

\textbf{Strategy.} Apply the mediant inequality at each step.  Since the mediant lies between the endpoints, replacing one endpoint preserves containment and makes the interval smaller.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is an induction-on-steps proof for an iterative construction.
Each stage is justified by the mediant inequality.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.49 and the order properties of real numbers.
\end{remark*}


\subsection*{Theorem 10.103 — Stern--Brocot Limit Theorem}
\label{prf:grep-stern-brocot-limit-theorem}

\begin{proof}
\Claim Every irrational number $x>0$ determines a unique infinite path in the Stern--Brocot tree.  The sequence of rational numbers encountered along this path converges to $x$.
Conversely, every infinite path in the Stern--Brocot tree determines a unique positive irrational number as its limit.

\textbf{Given.} An irrational number $x>0$, or alternatively an infinite path in the Stern--Brocot tree.

\textbf{Goal.} Prove both directions: existence/uniqueness of the path from $x$, and existence/uniqueness of the limit from the path.

\textbf{Strategy.} For the forward direction, repeatedly refine Stern--Brocot intervals according to whether $x$ lies left or right of the mediant.  For the reverse direction, show that an infinite path yields nested rational intervals whose lengths tend to $0$, and use this to define a unique real limit point.

% -------------------------------------------------------
% YOUR PROOF HERE
% -------------------------------------------------------

\end{proof}

\begin{remark*}[Proof shape]
This is the capstone structural proof of the chapter.  It is naturally divided
into two directions and is best approached through nested intervals whose
lengths shrink to zero.
\end{remark*}

\begin{remark*}[Dependencies]
Uses Theorem~10.96, the mediant-gap structure (especially Lemma~10.70), and a
nested-interval or completeness argument in $\mathbb{R}$.
\end{remark*}


\input{volume-ii/rationals/proofs/proof-stern-brocot}