% ==================================================
% Linearity
% Discrete Calculus Notes Stub
% ==================================================

\subsection{Properties of the Difference Operator}

The forward difference operator satisfies several algebraic properties
that closely resemble the familiar rules of differentiation in ordinary
calculus.

\subsection*{Linearity}

\begin{proposition}[Linearity]
For functions \(f,g : \mathbb{Z} \to \mathbb{R}\) and constants \(a,b\),

\[
\Delta (a f + b g)
=
a\,\Delta f + b\,\Delta g.
\]
\end{proposition}

\begin{proof}
By definition of the difference operator,

\[
\begin{aligned}
\Delta (a f + b g)(n)
&= (a f + b g)(n+1) - (a f + b g)(n) \\
&= a f(n+1) + b g(n+1) - a f(n) - b g(n) \\
&= a(f(n+1)-f(n)) + b(g(n+1)-g(n)) \\
&= a\,\Delta f(n) + b\,\Delta g(n).
\end{aligned}
\]
\end{proof}

\subsection*{Constant Rule}

\begin{proposition}
If \(c\) is a constant function, then

\[
\Delta c = 0.
\]
\end{proposition}

\begin{proof}
\[
\Delta c(n) = c - c = 0.
\]
\end{proof}

\begin{remark}
This mirrors the familiar rule from calculus:

\[
\frac{d}{dx} c = 0.
\]
\end{remark}

\subsection*{The Shift Operator}

\begin{definition}[Shift Operator]
The \emph{shift operator} \(E\) is defined by

\[
E f(n) = f(n+1).
\]
\end{definition}

\begin{remark}
The shift operator moves the input of the function forward by one step.
It plays a central role in the algebraic formulation of finite differences.
\end{remark}

\subsection*{Identity Operator}

\begin{definition}
The \emph{identity operator} \(I\) is defined by

\[
I f(n) = f(n).
\]
\end{definition}

\subsection*{Relationship Between Operators}

The difference operator can be expressed in terms of the shift operator.

\begin{theorem}
\[
\Delta = E - I.
\]
\end{theorem}

\begin{proof}
Applying the operator \(E-I\) to a function \(f\) gives

\[
(E-I)f(n) = Ef(n) - If(n).
\]

Using the definitions of \(E\) and \(I\),

\[
(E-I)f(n) = f(n+1) - f(n) = \Delta f(n).
\]

Thus

\[
\Delta = E - I.
\]
\end{proof}

\begin{remark}
This identity reveals an important algebraic structure underlying
discrete calculus.  Higher differences can now be expressed using the
binomial theorem:

\[
\Delta^n = (E-I)^n.
\]

This explains why binomial coefficients naturally appear in formulas
for higher finite differences.
\end{remark}

\begin{remark}
Expressing the difference operator in the form

\[
\Delta = E - I
\]

reveals that finite differences possess a simple algebraic structure.
The operator \(E\) shifts a function forward by one step,
while \(I\) leaves it unchanged.  The forward difference therefore
measures the discrepancy between a shifted function and the original one.

This formulation allows higher differences to be computed using ordinary
algebra.  Repeated application of the difference operator gives

\[
\Delta^n = (E-I)^n.
\]

Applying the binomial theorem to expand the right-hand side yields

\[
\Delta^n
=
\sum_{k=0}^{n}
(-1)^{\,n-k}
\binom{n}{k}
E^{k}.
\]

When this operator acts on a function \(f\), the powers of the shift
operator produce shifted values of the function:

\[
E^{k}f(n)=f(n+k).
\]

Consequently,

\[
\Delta^n f(n)
=
\sum_{k=0}^{n}
(-1)^{\,n-k}
\binom{n}{k}
f(n+k).
\]

Thus the pattern of coefficients in higher finite differences is governed
by the binomial coefficients.  This connection explains why Pascal's
triangle appears naturally in difference tables and why binomial
coefficients play a central role throughout discrete calculus.
\end{remark}


\subsection{Difference Tables and Polynomial Detection}

One of the most useful tools in discrete calculus is the \emph{difference
table}.  By repeatedly applying the difference operator to a sequence,
one can detect whether the sequence arises from a polynomial function and
determine its degree.

\begin{definition}[Difference Table]
Given a sequence \(f(0), f(1), f(2), \dots\), a \emph{difference table}
is constructed by listing the successive values of the function and then
computing their forward differences:

\[
\Delta f(n) = f(n+1)-f(n).
\]

The process may be repeated to obtain higher differences
\(\Delta^2 f\), \(\Delta^3 f\), and so forth.
\end{definition}

\subsection*{Example}

Consider the function

\[
f(n) = n^2.
\]

The values of the function and its successive differences are shown below.

\[
\begin{array}{c|c|c|c}
n & f(n) & \Delta f(n) & \Delta^2 f(n) \\
\hline
0 & 0  &   &   \\
1 & 1  & 1 &   \\
2 & 4  & 3 & 2 \\
3 & 9  & 5 & 2 \\
4 & 16 & 7 & 2 \\
5 & 25 & 9 & 2
\end{array}
\]

The first differences form the sequence

\[
1,3,5,7,9,\dots
\]

and the second differences are constant:

\[
2,2,2,\dots
\]

\begin{remark}
A constant second difference indicates that the original sequence is
generated by a polynomial of degree \(2\).
\end{remark}

\subsection*{General Principle}

\begin{theorem}[Polynomial Detection]
Let \(f(n)\) be a sequence generated by a polynomial of degree \(d\).
Then

\[
\Delta^d f(n)
\]

is constant, and

\[
\Delta^{d+1} f(n) = 0.
\]
\end{theorem}

\begin{remark}
This theorem is the discrete analogue of a familiar fact from calculus:
the \(d\)-th derivative of a degree \(d\) polynomial is constant, and the
next derivative is zero.
\end{remark}

\subsection*{Why This Matters}

Difference tables provide a practical method for identifying polynomial
sequences and determining their degree.  This observation leads to a
natural basis for polynomial functions in discrete calculus, known as the
\emph{falling factorial basis}.  In the next section we will see that these
functions interact particularly well with the difference operator and form
the foundation for Newton's discrete analogue of the Taylor expansion.

\begin{center}
\begin{tikzpicture}[
    every node/.style={font=\small},
    val/.style={},
    lbl/.style={font=\bfseries}
]

% Row labels
\node[lbl] at (-1.8, 0) {$f(n)$};
\node[lbl] at (-1.8,-1) {$\Delta f$};
\node[lbl] at (-1.8,-2) {$\Delta^2 f$};

% Function values for f(n)=n^2
\node[val] (f0) at (0,0) {$0$};
\node[val] (f1) at (1.6,0) {$1$};
\node[val] (f2) at (3.2,0) {$4$};
\node[val] (f3) at (4.8,0) {$9$};
\node[val] (f4) at (6.4,0) {$16$};
\node[val] (f5) at (8.0,0) {$25$};

% First differences
\node[val] (d0) at (0.8,-1) {$1$};
\node[val] (d1) at (2.4,-1) {$3$};
\node[val] (d2) at (4.0,-1) {$5$};
\node[val] (d3) at (5.6,-1) {$7$};
\node[val] (d4) at (7.2,-1) {$9$};

% Second differences
\node[val] (dd0) at (1.6,-2) {$2$};
\node[val] (dd1) at (3.2,-2) {$2$};
\node[val] (dd2) at (4.8,-2) {$2$};
\node[val] (dd3) at (6.4,-2) {$2$};

% Arrows showing subtraction pattern
\draw[->, thick] (f0) -- (d0);
\draw[->, thick] (f1) -- (d0);

\draw[->, thick] (f1) -- (d1);
\draw[->, thick] (f2) -- (d1);

\draw[->, thick] (f2) -- (d2);
\draw[->, thick] (f3) -- (d2);

\draw[->, thick] (f3) -- (d3);
\draw[->, thick] (f4) -- (d3);

\draw[->, thick] (f4) -- (d4);
\draw[->, thick] (f5) -- (d4);

\draw[->, thick] (d0) -- (dd0);
\draw[->, thick] (d1) -- (dd0);

\draw[->, thick] (d1) -- (dd1);
\draw[->, thick] (d2) -- (dd1);

\draw[->, thick] (d2) -- (dd2);
\draw[->, thick] (d3) -- (dd2);

\draw[->, thick] (d3) -- (dd3);
\draw[->, thick] (d4) -- (dd3);

\end{tikzpicture}
\end{center}

\begin{remark}
For the sequence \(f(n)=n^2\), the second differences are constant.
This indicates that the original sequence is generated by a polynomial
of degree \(2\). In general, if the \(k\)-th differences are constant,
then the sequence comes from a polynomial of degree \(k\).
\end{remark}

\begin{center}
\begin{tikzpicture}[
    every node/.style={font=\small},
    dot/.style={circle, fill=black, inner sep=1.2pt},
    lbl/.style={font=\bfseries}
]

% Labels
\node[lbl] at (-1.8, 0) {$f(n)$};
\node[lbl] at (-1.8,-1.2) {$\Delta f$};
\node[lbl] at (-1.8,-2.4) {$\Delta^2 f$};

% Top row
\node at (0,0) {$0$};
\node at (1.8,0) {$1$};
\node at (3.6,0) {$4$};
\node at (5.4,0) {$9$};
\node at (7.2,0) {$16$};
\node at (9.0,0) {$25$};

% First difference row
\node at (0.9,-1.2) {$1$};
\node at (2.7,-1.2) {$3$};
\node at (4.5,-1.2) {$5$};
\node at (6.3,-1.2) {$7$};
\node at (8.1,-1.2) {$9$};

% Second difference row
\node at (1.8,-2.4) {$2$};
\node at (3.6,-2.4) {$2$};
\node at (5.4,-2.4) {$2$};
\node at (7.2,-2.4) {$2$};

% Guide lines
\draw[gray] (0, -0.2) -- (0.9, -1.0);
\draw[gray] (1.8, -0.2) -- (0.9, -1.0);
\draw[gray] (1.8, -0.2) -- (2.7, -1.0);
\draw[gray] (3.6, -0.2) -- (2.7, -1.0);
\draw[gray] (3.6, -0.2) -- (4.5, -1.0);
\draw[gray] (5.4, -0.2) -- (4.5, -1.0);
\draw[gray] (5.4, -0.2) -- (6.3, -1.0);
\draw[gray] (7.2, -0.2) -- (6.3, -1.0);
\draw[gray] (7.2, -0.2) -- (8.1, -1.0);
\draw[gray] (9.0, -0.2) -- (8.1, -1.0);

\draw[gray] (0.9, -1.4) -- (1.8, -2.2);
\draw[gray] (2.7, -1.4) -- (1.8, -2.2);
\draw[gray] (2.7, -1.4) -- (3.6, -2.2);
\draw[gray] (4.5, -1.4) -- (3.6, -2.2);
\draw[gray] (4.5, -1.4) -- (5.4, -2.2);
\draw[gray] (6.3, -1.4) -- (5.4, -2.2);
\draw[gray] (6.3, -1.4) -- (7.2, -2.2);
\draw[gray] (8.1, -1.4) -- (7.2, -2.2);

\end{tikzpicture}
\end{center}





