% =========================================================
% Structural Induction
% =========================================================
\paragraph{Structural Induction (Preview).}

Induction on $\mathbb{N}$ is a special case of a more general
principle that applies to any recursively defined structure.

\begin{remark}[The general principle]
Whenever a set $X$ is defined by:
\begin{enumerate}
  \item a collection of \emph{base cases} (atomic elements), and
  \item a collection of \emph{constructor rules} (ways of building
    larger elements from smaller ones),
\end{enumerate}
structural induction is available: to prove $P(x)$ for all $x \in X$,
prove $P$ for all base cases, and prove that if $P$ holds for the
parts then $P$ holds for the whole.
\end{remark}

\begin{example}[Induction on algebraic expressions]
Define an \emph{algebraic term} over a ring $R$ by:
\begin{itemize}
  \item Base: any $r \in R$ is a term.
  \item Constructor: if $s, t$ are terms, so are $s + t$ and $s \cdot t$.
\end{itemize}
To prove a property $P$ holds for all terms:
\begin{enumerate}
  \item Prove $P(r)$ for all $r \in R$ (base cases).
  \item Prove: if $P(s)$ and $P(t)$ hold, then $P(s+t)$ holds.
  \item Prove: if $P(s)$ and $P(t)$ hold, then $P(s \cdot t)$ holds.
\end{enumerate}
This is structural induction on the recursive definition of terms.
\end{example}

\begin{remark}[Where structural induction appears in this project]
\begin{itemize}
  \item \textbf{Axiom Systems}: induction on $\mathbb{N}$ is proved
    from Peano P5, which is exactly the structural induction
    principle for $(\mathbb{N}, 0, S)$.
  \item \textbf{Abstract Algebra}: properties of polynomials over a ring
    are proved by induction on degree (structural induction on
    the recursive definition of polynomials).
  \item \textbf{Computer Science}: correctness of recursive algorithms
    is proved by structural induction on the input data structure
    (lists, trees, expressions).
\end{itemize}
The full development of structural induction appears in the
Abstract Algebra chapter.
\end{remark}
