% =========================================================
% Density of the Rational Numbers
% =========================================================

\subsection{Density of the Rational Numbers}
\label{sec:density-q}

\subsection{Definitions and Theorems}

Two different ideas are often confused when one first studies the rational numbers.

\begin{itemize}
  \item The \emph{Archimedean property} says that the natural numbers grow beyond every fixed
  number.
  \item The \emph{density of $\mathbb{Q}$} says that between any two distinct real numbers there is a
  rational number.
\end{itemize}

The first controls large-scale size; the second controls local refinement of intervals.

\begin{theorem}[Archimedean property of $\mathbb{Q}$]
\label{thm:archimedean-q}
For every $x\in\mathbb{Q}$ there exists $n\in\mathbb{N}$ such that
\[
n>x.
\]
Equivalently, for every $x>0$ there exists $n\in\mathbb{N}$ such that
\[
\frac{1}{n}<x.
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:archimedean-q]{Go to proof.}
\end{remark*}

\begin{remark}
This theorem says that the reciprocals of natural numbers become arbitrarily small.  It rules out the
existence of positive infinitesimals in $\mathbb{Q}$.
\end{remark}

\begin{theorem}[Density of $\mathbb{Q}$ in $\mathbb{R}$]
\label{thm:q-dense-in-r}
If $x,y\in\mathbb{R}$ and $x<y$, then there exists $q\in\mathbb{Q}$ such that
\[
x<q<y.
\]
\end{theorem}

\begin{remark*}[Dependency ceiling note]
This theorem is stated for $x, y \in \mathbb{R}$, and its proof invokes the
Archimedean property of $\mathbb{R}$ together with the well-ordering of
$\mathbb{Z}$.  At this point in the notes $\mathbb{R}$ has not yet been
constructed.  The theorem is placed here because its content concerns
$\mathbb{Q}$, but its formal proof belongs to the reals chapter.  The
student should treat it as a forward reference until
\hyperref[sec:reals]{the real numbers chapter} is complete.
\end{remark*}

\begin{remark*}[Proof]
\hyperref[prf:q-dense-in-r]{Go to proof.}
\end{remark*}

\begin{corollary}
\label{cor:infinitely-many-rationals-between}
Between any two distinct real numbers there are infinitely many rational numbers.
\end{corollary}

\begin{remark*}[Proof]
\hyperref[prf:infinitely-many-rationals-between]{Go to proof.}
\end{remark*}

\subsection{Consequences}

\begin{remark}
Density is one of the reasons the rational numbers are so useful in analysis.  Even when a number is
not rational, it can still be approximated arbitrarily well by rational numbers.
\end{remark}
