% =========================================================
% Linear Algebra — Structural Roadmap
% Driver Text: Sheldon Axler, Linear Algebra Done Right
% File: 100-linear-algebra/linear-algebra.tex
% =========================================================

\subsection{Linear Algebra}

% =========================================================
% Structural Roadmap
% =========================================================
\subsubsection*{Structural Roadmap}

The development of linear algebra in this project follows
the definition--theorem--structure architecture used
throughout the analysis volumes.

The primary driver is \textit{Linear Algebra Done Right}
by Sheldon Axler. The emphasis is on linear maps,
invariant structure, and conceptual clarity
rather than computational techniques.

Each major topic is organized as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems $\longrightarrow$ Consequences and Structural Insight}
\end{center}

The global progression is:

\begin{enumerate}
  \item Vector spaces and subspaces
  \item Linear combinations, span, and linear independence
  \item Bases and dimension
  \item Linear maps
  \item Null space and range
  \item Matrix representations of linear maps
  \item Eigenvalues and eigenvectors
  \item Invariant subspaces
  \item Triangularization and diagonalization
  \item Inner product spaces
  \item Orthogonality and projections
  \item Self-adjoint and normal operators
  \item Spectral theorem (finite-dimensional case)
\end{enumerate}

\vspace{1em}

\begin{remark}
This treatment prioritizes structural understanding over
algorithmic manipulation. Determinants are introduced only
after linear maps and eigenvalues have been conceptually
understood.
\end{remark}

\begin{remark}
The central object of study is not matrices,
but linear maps between vector spaces.
Matrices serve only as representations relative to a basis.
\end{remark}

\begin{remark}[Structural Position]
Linear algebra is developed independently of analysis.
Its structural viewpoint will later support operator theory,
Hilbert spaces, and functional analysis.
\end{remark}

% =========================================================
% I. Core Linear Structure
% =========================================================

\input{linear-algebra/notes/notes-vector-spaces}

% =========================================================
% II. Linear Maps
% =========================================================

% \input{linear-algebra/notes/notes-linear-independence}
% \input{linear-algebra/notes/notes-basis-dimension}
% \input{linear-algebra/notes/notes-linear-maps}
% \input{linear-algebra/notes/notes-null-range}
% \input{linear-algebra/notes/notes-matrix-representations}

% =========================================================
% III. Eigenstructure
% =========================================================

% \input{linear-algebra/notes/notes-eigenvalues}
% \input{linear-algebra/notes/notes-invariant-subspaces}
% \input{linear-algebra/notes/notes-diagonalization}

% =========================================================
% IV. Inner Product Structure
% =========================================================

% \input{linear-algebra/notes/notes-inner-products}
% \input{linear-algebra/notes/notes-orthogonality}
% \input{linear-algebra/notes/notes-spectral-theorem}

% =========================================================
% Future Sections (To Be Developed)
% =========================================================

% \input{linear-algebra/notes/notes-determinants}
% \input{linear-algebra/notes/notes-jordan-form}
% \input{linear-algebra/notes/notes-advanced-operators}

% =========================================================
% Logical Positioning
% =========================================================

