% =========================================================
\subsubsection{Balls in Metric Spaces}
\label{sec:balls-metric-spaces}
% =========================================================

% ---- Toolkit ----
\begin{tcolorbox}[colback=gray!6, colframe=gray!40, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Balls in Metric Spaces --- Quick Reference}},
  fonttitle=\small\bfseries]
\small
\begin{tabular}{l l l l}
\toprule
\textbf{Label} & \textbf{Statement} & \textbf{Proof method} & \textbf{Detail} \\
\midrule
D\ref{def:open-ball}
  & $B_r(x) := \{\,y\in X : d(x,y) < r\,\}$
  & Definition
  & \hyperref[def:open-ball]{$\downarrow$} \\
D\ref{def:closed-ball}
  & $\overline{B}_r(x) := \{\,y\in X : d(x,y) \le r\,\}$
  & Definition
  & \hyperref[def:closed-ball]{$\downarrow$} \\
D\ref{def:open-closed-set}
  & Open: $\forall x\in S,\;\exists r>0,\;B_r(x)\subseteq S$
  & Definition
  & \hyperref[def:open-closed-set]{$\downarrow$} \\
P\ref{prop:balls-open-closed}
  & $B_r(x)$ is open;\; $\overline{B}_r(x)$ is closed
  & Triangle inequality
  & \hyperref[prop:balls-open-closed]{$\downarrow$} \\
\bottomrule
\end{tabular}
\end{tcolorbox}

% ---- Open Ball ----
\begin{tcolorbox}[colback=propbox, colframe=propborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Definition (Open Ball)}},
  fonttitle=\small\bfseries]
\begin{definition}[Open Ball]
\label{def:open-ball}
Let $(X,d)$ be a metric space.
For $x \in X$ and $r > 0$, the \emph{open ball} centred at $x$ with radius $r$ is
\[
B_r(x)
:=
\{\, y \in X : d(x,y) < r \,\}.
\]
\end{definition}
\end{tcolorbox}

\begin{remark*}[Fully quantified form]
$\forall (X,d)\ \text{metric space},\;
 \forall x \in X,\;
 \forall r \in \mathbb{R}^{+},\quad
 B_r(x) := \{\, y \in X : d(x,y) < r \,\}.$
\end{remark*}

\input{volume-iii/analysis/metric-spaces/notes/balls/figure-open-ball-contains-open-balls}

\begin{remark*}[Intuition]
The open ball is the metric-space generalisation of an open interval in $\mathbb{R}$.
In $(\mathbb{R}, |\cdot|)$ it recovers the interval $(x-r, x+r)$.
In $(\mathbb{R}^2, \|\cdot\|_2)$ it is the open disc of radius $r$.
The strict inequality $d(x,y) < r$ means the boundary circle is excluded.
\end{remark*}
\begin{remark*}
Every point of an open ball has positive distance from the boundary of the ball.
This allows a smaller ball to be placed around any interior point.
\end{remark*}

% ---- Closed Ball ----
\begin{tcolorbox}[colback=propbox, colframe=propborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Definition (Closed Ball)}},
  fonttitle=\small\bfseries]
\begin{definition}[Closed Ball]
\label{def:closed-ball}
Let $(X,d)$ be a metric space.
For $x \in X$ and $r > 0$, the \emph{closed ball} centred at $x$ with radius $r$ is
\[
\overline{B}_r(x)
:=
\{\, y \in X : d(x,y) \le r \,\}.
\]
\end{definition}
\end{tcolorbox}

\begin{remark*}[Fully quantified form]
$\forall (X,d)\ \text{metric space},\;
 \forall x \in X,\;
 \forall r \in \mathbb{R}^{+},\quad
 \overline{B}_r(x) := \{\, y \in X : d(x,y) \le r \,\}.$
\end{remark*}

\input{volume-iii/analysis/metric-spaces/notes/balls/figure-closed-open-balls}

\begin{remark*}[Common error]
The closure of the open ball $\overline{B_r(x)}$ need not equal the closed
ball $\overline{B}_r(x)$ in a general metric space.
In $\mathbb{R}^n$ with the standard metric they agree, but in a discrete metric
space $B_r(x) = \{x\}$ for $r \le 1$ and $\overline{B_r(x)} = \{x\}$ while
$\overline{B}_r(x) = X$ for $r = 1$.
The notation is therefore intentionally distinct.
\end{remark*}

\input{volume-iii/analysis/metric-spaces/notes/balls/figure-open-closed-balls}

\begin{remark*}[Interior points of a ball]
Every point inside a ball that is not on the boundary admits a smaller
open ball around it.

Let $(X,d)$ be a metric space and let $y \in B_r(x)$, so that
\[
d(x,y) < r.
\]
Define
\[
\varepsilon := r - d(x,y) > 0.
\]
Then
\[
B_{\varepsilon}(y) \subseteq B_r(x).
\]

Geometrically, the quantity $r - d(x,y)$ represents the remaining
distance from $y$ to the boundary of the ball $B_r(x)$.
Thus a smaller open ball centered at $y$ fits entirely inside $B_r(x)$.
\end{remark*}

\begin{remark}[Intervals generalise to balls]
In $(\mathbb{R},|\cdot|)$, open balls coincide with open intervals and
closed balls coincide with closed intervals:

\[
B_r(x) = (x-r,x+r), \qquad \overline{B}_r(x) = [x-r,x+r].
\]

Thus open and closed intervals on the real line generalise naturally
to open and closed balls in metric spaces.
\end{remark}



% ---- Open Set and Closed Set ----
\begin{tcolorbox}[colback=propbox, colframe=propborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Definition (Open Set, Closed Set)}},
  fonttitle=\small\bfseries]
\begin{definition}[Open Set, Closed Set]
\label{def:open-closed-set}
Let $(X, d)$ be a metric space and $S \subseteq X$.
\begin{itemize}
  \item $S$ is \emph{open} (in $X$) if for every $x \in S$ there exists
        $r > 0$ such that $B_r(x) \subseteq S$.
  \item $S$ is \emph{closed} (in $X$) if its complement $X \setminus S$
        is open.
\end{itemize}
\end{definition}
\end{tcolorbox}

\begin{remark*}[Fully quantified form]
\begin{align*}
S\ \text{open}
  &\iff \forall x \in S,\;
         \exists r \in \mathbb{R}^{+},\;
         B_r(x) \subseteq S. \\
S\ \text{closed}
  &\iff X \setminus S\ \text{is open} \\
  &\iff \forall x \in X \setminus S,\;
         \exists r \in \mathbb{R}^{+},\;
         B_r(x) \subseteq X \setminus S.
\end{align*}
\end{remark*}

\begin{remark*}[Common error]
Open and closed are \emph{not} opposites.
A set can be both open and closed (e.g.\ $\emptyset$ and $X$ itself),
or neither (e.g.\ the half-open interval $[0,1)$ in $\mathbb{R}$).
\end{remark*}

\input{volume-iii/analysis/metric-spaces/notes/balls/figure-open-set}

% ---- Balls are open and closed ----
\begin{proposition}[\texorpdfstring{\hyperref[prf:balls-open-closed]{Balls are Open and Closed}}{Balls are Open and Closed}]
\label{prop:balls-open-closed}
Let $(X, d)$ be a metric space, $x \in X$, and $r > 0$.
\begin{enumerate}[label=(\roman*)]
  \item $B_r(x)$ is an open set.
  \item $\overline{B}_r(x)$ is a closed set.
\end{enumerate}
\end{proposition}

\begin{remark*}[Fully quantified form]
\begin{align*}
&\forall (X,d)\ \text{metric space},\;
 \forall x \in X,\;
 \forall r \in \mathbb{R}^{+}: \\
&\quad \text{(i)}\ \forall y \in B_r(x),\;
       \exists \alpha \in \mathbb{R}^{+},\;
       B_\alpha(y) \subseteq B_r(x). \\
&\quad \text{(ii)}\ X \setminus \overline{B}_r(x)\ \text{is open},
       \text{ i.e.}\ \forall y \notin \overline{B}_r(x),\;
       \exists \alpha \in \mathbb{R}^{+},\;
       B_\alpha(y) \cap \overline{B}_r(x) = \emptyset.
\end{align*}
\end{remark*}

\begin{remark*}[Intuition]
For~(i): given $y \in B_r(x)$, the slack $\alpha := r - d(x,y) > 0$ is
the room left before the boundary of $B_r(x)$.
The triangle inequality then guarantees $B_\alpha(y) \subseteq B_r(x)$.
For~(ii): the same idea with $\alpha := d(x,y) - r > 0$ applied to a
point outside the closed ball.
\end{remark*}

\input{volume-iii/analysis/metric-spaces/notes/balls/figure-slack-ball}

\begin{proposition}[Trivial open and closed sets and balls]
Let $(X,d)$ be a metric space.

\begin{enumerate}
\item The empty set $\varnothing$ and the whole space $X$ are open.
\item The empty set $\varnothing$ and the whole space $X$ are closed.
\item Every open ball $B_r(x)$ is an open set.
\item Every closed ball $\overline{B}_r(x)=\{y\in X : d(x,y)\le r\}$ is a closed set.
\end{enumerate}
\end{proposition}

\begin{proof}
(1) The empty set is open because the defining condition for openness
\[
\forall x \in U,\;\exists \varepsilon>0 \text{ such that } B_\varepsilon(x)\subseteq U
\]
is vacuously true when $U=\varnothing$.

The set $X$ is open because for any $x\in X$ and any $\varepsilon>0$,
\[
B_\varepsilon(x)\subseteq X.
\]

(2) A set is closed if its complement is open. Since
\[
X\setminus \varnothing = X
\qquad\text{and}\qquad
X\setminus X = \varnothing,
\]
and both sets are open by part (1), it follows that $\varnothing$ and $X$ are closed.

(3) Let $x_0\in B_r(x)$. Then $d(x,x_0)<r$. Choose $\varepsilon>0$ such that
\[
\varepsilon < r-d(x,x_0).
\]
If $y\in B_\varepsilon(x_0)$ then $d(x_0,y)<\varepsilon$, and by the triangle
inequality
\[
d(x,y)\le d(x,x_0)+d(x_0,y) < d(x,x_0)+\varepsilon < r.
\]
Thus $y\in B_r(x)$ and therefore $B_\varepsilon(x_0)\subseteq B_r(x)$.
Hence $B_r(x)$ is open.

(4) Let $y\notin \overline{B}_r(x)$. Then $d(x,y)>r$. Set
\[
\varepsilon = d(x,y)-r>0.
\]
If $z\in B_\varepsilon(y)$ then $d(y,z)<\varepsilon$, and the reverse triangle
inequality gives
\[
d(x,z)\ge d(x,y)-d(y,z) > r.
\]
Thus $z\notin \overline{B}_r(x)$, so
\[
B_\varepsilon(y)\subseteq X\setminus \overline{B}_r(x).
\]
Hence the complement of $\overline{B}_r(x)$ is open, and therefore
$\overline{B}_r(x)$ is closed.
\end{proof}