% =========================================================
% Rational Numbers
% =========================================================

\section{Rational Numbers}
\label{sec:rational-numbers}

\subsection{Definitions and Theorems}

The rational numbers are built from the integers in essentially the same way that the integers were
built from the natural numbers: one begins with concrete representatives, identifies those that ought
to represent the same object, and then passes to equivalence classes.

\begin{definition}[Fraction pairs]
\label{def:fraction-pairs}

Let $\mathbb{Z}$ denote the set of integers and define
\[
W = \mathbb{Z} \times (\mathbb{Z}\setminus\{0\}).
\]
Thus
\[
W = \{(a,b) : a,b\in\mathbb{Z},\ b\neq 0\}.
\]

Each ordered pair $(a,b)$ is intended to represent the formal fraction $a/b$.

\end{definition}

\begin{remark}[Numerator and denominator]
\label{rem:numerator-denominator}

If $(a,b)\in W$, then $a$ is called the \emph{numerator} and $b$ is called the
\emph{denominator}.  At this stage $(a,b)$ is still only an ordered pair; it is not yet a rational
number.

\end{remark}

\begin{remark}
Different pairs should represent the same rational number.  For example,
\[
(1,2),\quad (2,4),\quad (3,6)
\]
all ought to describe the same quantity.  To make this precise we introduce an equivalence relation.
\end{remark}

\begin{definition}[Equality of fraction pairs]
\label{def:fraction-pair-equality}

For $(a,b),(c,d)\in W$, define
\[
(a,b) \sim (c,d)
\quad\Longleftrightarrow\quad
ad = bc.
\]

In fully quantified form,
\[
\forall (a,b),(c,d)\in W,
\qquad
(a,b)\sim(c,d)
\Longleftrightarrow
ad=bc.
\]

\end{definition}

\begin{remark}
This is the familiar cross-multiplication test for equality of fractions.  For example,
\[
(1,2)\sim(2,4)
\]
because
\[
1\cdot 4 = 2\cdot 2.
\]
\end{remark}

\begin{theorem}
\label{thm:fraction-pair-equivalence}
The relation $\sim$ is an equivalence relation on $W$.
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:fraction-pair-equivalence]{Go to proof.}
\end{remark*}

\begin{remark}
Once this is known, the equivalence classes of $\sim$ partition $W$.  Each class is what we shall
call a rational number.
\end{remark}

\begin{definition}[Rational numbers]
\label{def:rational-numbers}

The set of \emph{rational numbers} is the quotient
\[
\mathbb{Q} = W/\sim.
\]
If $(a,b)\in W$, its equivalence class is denoted by
\[
[(a,b)].
\]
We interpret $[(a,b)]$ as the rational number represented informally by $a/b$.

\end{definition}

\begin{definition}[Distinguished elements]
\label{def:rational-distinguished}

Define
\[
0 := [(0,1)]
\qquad\text{and}\qquad
1 := [(1,1)].
\]

\end{definition}

\begin{remark}
These definitions are well-defined because
\[
(0,1)\sim(0,b) \quad (b\neq 0)
\qquad\text{and}\qquad
(1,1)\sim(a,a) \quad (a\neq 0).
\]
\end{remark}

\begin{definition}[Addition and multiplication]
\label{def:rational-operations}

Let $[(a,b)],[(c,d)]\in\mathbb{Q}$, where $(a,b),(c,d)\in W$.  Define
\[
[(a,b)] + [(c,d)] := [(ad+bc,bd)]
\]
and
\[
[(a,b)]\cdot[(c,d)] := [(ac,bd)].
\]

\end{definition}

\begin{lemma}
\label{lem:rational-addition-well-defined}
Let $a,b,c,d,e,f,g,h\in\mathbb{Z}$ with $bdfh\neq 0$. If
\[
(a,b)\sim(e,f)
\qquad\text{and}\qquad
(c,d)\sim(g,h),
\]
then
\[
(ad+bc,bd)\sim(eh+fg,fh).
\]
\end{lemma}

\begin{remark*}[Proof]
\hyperref[prf:rational-addition-well-defined]{Go to proof.}
\end{remark*}

\begin{lemma}
\label{lem:rational-multiplication-well-defined}
Let $a,b,c,d,e,f,g,h\in\mathbb{Z}$ with $bdfh\neq 0$. If
\[
(a,b)\sim(e,f)
\qquad\text{and}\qquad
(c,d)\sim(g,h),
\]
then
\[
(ac,bd)\sim(eg,fh).
\]
\end{lemma}

\begin{remark*}[Proof]
\hyperref[prf:rational-multiplication-well-defined]{Go to proof.}
\end{remark*}

\begin{remark}[Why well-definedness matters]
\label{rem:well-definedness-matters}

These formulas are written in terms of representatives.  Because one rational number has many
representatives, it must be proved that choosing different representatives leads to the same
resulting equivalence class.  This is the first serious instance of a very common pattern in
mathematics:
\[
\text{representatives}
\longrightarrow
\text{operations}
\longrightarrow
\text{well-defined operations on equivalence classes}.
\]

\end{remark}

\begin{remark}[Three levels of representation]
\label{rem:rational-three-levels}

When working with rational numbers, it is essential to distinguish among the following three objects.

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Object} & \textbf{Lives in} & \textbf{Meaning} \\
\midrule
$(a,b)$ & $W$ & raw integer pair \\
$[(a,b)]$ & $\mathbb{Q}$ & rational number \\
$\dfrac{a}{b}$ & notation & informal representation of $[(a,b)]$ \\
\bottomrule
\end{tabular}
\end{center}

The pair $(a,b)$ is not itself a rational number.  The rational number is the entire equivalence
class
\[
[(a,b)] = \{(c,d)\in W : ad = bc\}.
\]

\end{remark}

\begin{remark}[A rational number is an infinite set]
\label{rem:rational-infinite-set}

For every nonzero integer $k$,
\[
(ka,kb)\sim(a,b).
\]
Hence the class $[(a,b)]$ contains infinitely many representatives:
\[
(a,b),(2a,2b),(3a,3b),\dots,(-a,-b),(-2a,-2b),\dots
\]
Thus every rational number is an infinite set.
\end{remark}

\begin{remark}[Canonical representative]
\label{rem:reduced-rational}

Although a rational number has infinitely many representatives, each nonzero rational number can be
written uniquely in the form
\[
\frac{a}{b}
\qquad\text{with}\qquad
\gcd(a,b)=1,
\quad b>0.
\]
This is the \emph{reduced form} of the rational number.

\end{remark}

\subsection{Consequences}

\begin{remark}[Quotient structures]
\label{rem:quotient-pattern}

The construction of $\mathbb{Q}$ is the first major example in these notes of the general pattern
\[
\text{set} + \text{equivalence relation} \longrightarrow \text{quotient structure}.
\]
The same pattern later appears in modular arithmetic, quotient groups, quotient rings, quotient
spaces, and projective geometry.

\end{remark}

\subsection{Algebra of Equivalence Classes}

\begin{remark}[Role of the equivalence-class arithmetic theorems]
\label{rem:ec-arithmetic-role}

The theorems in this subsection work entirely at the level of
equivalence classes and integer arithmetic.  They require nothing
beyond \hyperref[def:rational-operations]{Definition~\ref*{def:rational-operations}},
the well-definedness lemmas, and facts about $\mathbb{Z}$.

In particular, \hyperref[thm:ec-negation]{Theorem~\ref*{thm:ec-negation}}
and \hyperref[thm:ec-reciprocal]{Theorem~\ref*{thm:ec-reciprocal}} do
not appeal to field axioms.  They \emph{establish}, by direct
computation, that $[(-a,b)]$ and $[(b,a)]$ satisfy the defining
conditions for additive and multiplicative inverses respectively.
These facts are therefore \emph{inputs to} the proof that $\mathbb{Q}$
is a field, not consequences of it.  The field proof can cite them
rather than repeat the computation.

The later \hyperref[sec:rational-algebra]{rational algebra section}
re-expresses the same identities using fraction notation $a/b$ inside
the already-established field.  Those re-derivations are about
\emph{notation and usability}, not about new mathematical content.

\end{remark}

\begin{theorem}[Integer embedding preserves addition]
\label{thm:ec-integer-embedding-add}
For all $m,n\in\mathbb{Z}$,
\[
[(m,1)] + [(n,1)] = [(m+n,\,1)].
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-integer-embedding-add]{Go to proof.}
\end{remark*}

\begin{theorem}[Integer embedding preserves multiplication]
\label{thm:ec-integer-embedding-mul}
For all $m,n\in\mathbb{Z}$,
\[
[(m,1)] \cdot [(n,1)] = [(mn,\,1)].
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-integer-embedding-mul]{Go to proof.}
\end{remark*}

\begin{remark}[The integer embedding]
\label{rem:integer-embedding}
These two theorems establish that the map $n\mapsto[(n,1)]$ is a ring
homomorphism from $\mathbb{Z}$ into $\mathbb{Q}$: it sends the integer
operations to the rational operations.  Moreover $[(m,1)]=[(n,1)]$ if
and only if $m=n$, so the map is injective and $\mathbb{Z}$ sits inside
$\mathbb{Q}$ as a subring.  We therefore identify each integer $n$ with
the rational class $[(n,1)]$ and write $n$ for $[(n,1)]$ without further
comment.
\end{remark}

\begin{theorem}[Zero class criterion]
\label{thm:ec-zero-numerator}
For every $[(a,b)]\in\mathbb{Q}$,
\[
[(a,b)] = [(0,1)]
\quad\Longleftrightarrow\quad
a = 0.
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-zero-numerator]{Go to proof.}
\end{remark*}

\begin{theorem}[Scaling of equivalence classes]
\label{thm:ec-scaling}
For every $[(a,b)]\in\mathbb{Q}$ and every nonzero integer $e\neq 0$,
\[
[(a,b)] = [(ae,be)].
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-scaling]{Go to proof.}
\end{remark*}

\begin{theorem}[Negation of an equivalence class]
\label{thm:ec-negation}
For every $[(a,b)]\in\mathbb{Q}$,
\[
-[(a,b)] = [(-a,b)] = [(a,-b)].
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-negation]{Go to proof.}
\end{remark*}

\begin{theorem}[Subtraction of equivalence classes]
\label{thm:ec-subtraction}
For every $[(a,b)],[(c,d)]\in\mathbb{Q}$,
\[
[(a,b)] - [(c,d)] = [(ad-bc,\,bd)].
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-subtraction]{Go to proof.}
\end{remark*}

\begin{theorem}[Reciprocal of a nonzero equivalence class]
\label{thm:ec-reciprocal}
For every $[(a,b)]\in\mathbb{Q}$ with $a\neq 0$,
\[
[(a,b)]^{-1} = [(b,a)].
\]
\end{theorem}

\begin{remark*}[Proof]
\hyperref[prf:ec-reciprocal]{Go to proof.}
\end{remark*}
