% =========================================================
% Bounded Sequences and Types of Bounds
% =========================================================

\subsection{Bounded Sequences and Types of Bounds}

\subsubsection{Bounded Sequences}

% ---------------------------------------------------------
\subsubsection{Basic Definitions}

\begin{definition}[Bounded sequence]
A sequence $(x_n)$ in $\mathbb{R}$ is called \emph{bounded} if there exists a real
number $M>0$ such that
\[
|x_n| \le M
\quad\text{for all } n\in\mathbb{N}.
\]
\end{definition}

\begin{remark}[Equivalent form]
A sequence $(x_n)$ is bounded if and only if there exist real numbers
$m$ and $M$ such that
\[
m \le x_n \le M
\quad\text{for all } n\in\mathbb{N}.
\]
\end{remark}

\begin{remark}[Logical form]
\[
\exists M>0\;
\forall n\in\mathbb{N}\;
\bigl(|x_n|\le M\bigr).
\]
\[
\exists m,M\in\mathbb{R}\;
\forall n\in\mathbb{N}\;
\bigl(m\le x_n \le M\bigr).
\]
\end{remark}

\begin{remark}[Functional viewpoint]
A sequence $\{x_n\}_{n=1}^\infty$ is bounded precisely when the function
\[
x : \mathbb{N} \to \mathbb{R}
\]
defined by $x(n)=x_n$ is bounded on its domain.
\end{remark}

\begin{remark}
A sequence $\{x_n\}_{n=1}^\infty$ in $\mathbb{R}$ is bounded
if and only if the subset
\[
\{x_n : n \in \mathbb{N}\} \subseteq \mathbb{R}
\]
is bounded.
\end{remark}

\begin{remark}[Three structural viewpoints of a sequence]
A sequence may be understood at three distinct but equivalent levels:

\begin{enumerate}
    \item \textbf{Function viewpoint:}
    A sequence $\{x_n\}_{n=1}^\infty$ is a function
    \[
    x : \mathbb{N} \to \mathbb{R}, \quad x(n)=x_n.
    \]

    \item \textbf{Set (range) viewpoint:}
    The values of the sequence form the subset
    \[
    \{x_n : n \in \mathbb{N}\} \subseteq \mathbb{R},
    \]
    which is the range of the function.

    \item \textbf{Order-theoretic viewpoint:}
    Boundedness of the sequence is not a property of the index set
    $\mathbb{N}$ but of this subset of $\mathbb{R}$.
    Thus the sequence is bounded if and only if its range
    $\{x_n : n \in \mathbb{N}\}$ is bounded in $\mathbb{R}$.
\end{enumerate}

These perspectives are logically equivalent but conceptually distinct,
and each becomes useful in different parts of analysis.
\end{remark}



% ---------------------------------------------------------
% Types of Bounds for Sequences
% ---------------------------------------------------------

\begin{definition}[Upper bound]
A real number $M$ is an \emph{upper bound} for the sequence $(x_n)$ if
\[
x_n \le M
\quad\text{for all } n\in\mathbb{N}.
\]
\end{definition}

\begin{definition}[Lower bound]
A real number $m$ is a \emph{lower bound} for the sequence $(x_n)$ if
\[
m \le x_n
\quad\text{for all } n\in\mathbb{N}.
\]
\end{definition}

\begin{definition}[Bounded above / below]
A sequence $(x_n)$ is said to be
\begin{itemize}
\item \emph{bounded above} if it has at least one upper bound;
\item \emph{bounded below} if it has at least one lower bound;
\item \emph{bounded} if it is both bounded above and bounded below.
\end{itemize}
\end{definition}

% ---------------------------------------------------------
% Least Upper and Greatest Lower Bounds
% ---------------------------------------------------------

\begin{definition}[Supremum of a sequence]
If the set
\[
\{x_n : n\in\mathbb{N}\}
\]
is bounded above, its least upper bound is called the \emph{supremum}
of the sequence and is denoted
\[
\sup\{x_n\}
\quad\text{or}\quad
\sup_{n\in\mathbb{N}} x_n.
\]
\end{definition}

\begin{definition}[Infimum of a sequence]
If the set
\[
\{x_n : n\in\mathbb{N}\}
\]
is bounded below, its greatest lower bound is called the \emph{infimum}
of the sequence and is denoted
\[
\inf\{x_n\}
\quad\text{or}\quad
\inf_{n\in\mathbb{N}} x_n.
\]
\end{definition}

\begin{remark}
The supremum or infimum of a sequence need not be attained by any term of the sequence.
\end{remark}

% ---------------------------------------------------------
% Extremal Values
% ---------------------------------------------------------

\begin{definition}[Maximum and minimum of a sequence]
A sequence $(x_n)$ has a \emph{maximum} if there exists $N\in\mathbb{N}$ such that
\[
x_N \ge x_n
\quad\text{for all } n\in\mathbb{N}.
\]
A sequence $(x_n)$ has a \emph{minimum} if there exists $N\in\mathbb{N}$ such that
\[
x_N \le x_n
\quad\text{for all } n\in\mathbb{N}.
\]
\end{definition}

\begin{remark}
If a maximum exists, then
\[
\max\{x_n\} = \sup\{x_n\},
\]
and similarly for a minimum and the infimum.
\end{remark}

% ---------------------------------------------------------
\subsubsection{Main Theorems}

% (No independent theorems were present in the source extract.)
% This section is definitional in nature.

% ---------------------------------------------------------
\subsubsection{Consequences}

The logical implication of this entire section is:

\[
\text{Bounded sequence}
\Longleftrightarrow
\text{Range } \{x_n\} \text{ is bounded in } \mathbb{R}.
\]

Thus sequence boundedness reduces to set boundedness.

Moreover:

\[
\text{Maximum} \Rightarrow \text{Supremum attained},
\]
\[
\text{Minimum} \Rightarrow \text{Infimum attained}.
\]

But the converse need not hold unless the supremum or infimum
is actually realized by some term of the sequence.

\begin{remark}[Logical Structure]
The structural progression is:

\[
\text{Upper/Lower bounds}
\Rightarrow
\text{Boundedness}
\Rightarrow
\text{Supremum/Infimum}
\Rightarrow
\text{Maximum/Minimum}.
\]

All later extremal and limit arguments depend on this hierarchy.
\end{remark}
