% =========================================================
% Proof: Closed iff Complement Open
% Mode: standard
% =========================================================

\subsection*{Closed Sets and Open Complements}
\label{prf:closed-iff-complement-open}

\begin{remark*}[Return]
$\leftarrow$ \hyperref[thm:closed-iff-complement-open]{Back to Theorem
(Closed iff Complement Open) in Notes}
\end{remark*}

\begin{proof}
\Claim Let $(X,d)$ be a metric space and $F \subseteq X$.
$F$ is closed (i.e.\ $F$ contains all limits of sequences in $F$)
if and only if $X \setminus F$ is open.

\Given A metric space $(X,d)$ and $F \subseteq X$.

\Goal To prove $F$ closed $\Longleftrightarrow$ $X \setminus F$ open.

\Strategy Both directions proceed by contraposition at the level of points:
a point outside a closed set has a ball keeping it outside;
a sequence in $F$ heading for the exterior contradicts closedness.

\medskip
\noindent\textbf{($\Rightarrow$) $F$ closed $\Rightarrow$ $X \setminus F$ open.}

\Let $y \in X \setminus F$. We must find $r > 0$ with $B_r(y) \subseteq X \setminus F$.

Suppose for contradiction that no such $r$ exists. Then for each $n \ge 1$,
$B_{1/n}(y)$ contains a point $x_n \in F$.
The sequence $(x_n)$ lies in $F$ and $d(x_n, y) < 1/n \to 0$, so $x_n \to y$.

\ByAss $F$ is closed: since $(x_n) \subseteq F$ and $x_n \to y$, we have $y \in F$.
This contradicts $y \in X \setminus F$. \Therefore the $r$ exists and $X \setminus F$ is open.

\medskip
\noindent\textbf{($\Leftarrow$) $X \setminus F$ open $\Rightarrow$ $F$ closed.}

\Let $(x_n) \subseteq F$ with $x_n \to x$.

Suppose for contradiction that $x \notin F$, so $x \in X \setminus F$.

\ByAss $X \setminus F$ is open: $\exists r > 0$ with $B_r(x) \subseteq X \setminus F$.

\ByAss $x_n \to x$: $\exists N$ with $n \ge N \Rightarrow d(x_n, x) < r$, i.e.\ $x_n \in B_r(x) \subseteq X \setminus F$.

But $x_n \in F$, so $x_n \in F \cap (X \setminus F) = \varnothing$. Contradiction.

\Therefore $x \in F$. Since the sequence $(x_n) \subseteq F$ was arbitrary, $F$ is closed. \AsReq
\end{proof}

\begin{remark*}[Proof shape]
Both directions use a sequence argument: build a sequence converging
to the relevant point and derive a contradiction if the conclusion fails.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item Definition~\ref{def:open-closed-set}: $F$ closed means $F$ is sequentially closed.
  \item Definition~\ref{def:open-closed-set}: $U$ open means every point has a ball inside.
  \item Archimedean property: $1/n \to 0$.
\end{itemize}
\end{remark*}
